--- 
title: "Data Analytics"
author: "Shivam Upadhyay"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: book
bibliography: [book.bib, packages.bib]
description: |
  This is a minimal example.
link-citations: yes
url: 'https\://bookdown.org/shivam7898/xADSM'
github-repo: "shivam7898/xADSM"
subtitle: "ADSM: Statistics and R"
---

# About {#FIRST .unnumbered .tabset .tabset-fade} 
[Goto LAST](#LAST "End of the Document")

~~"Copy/Paste is the mother of learning."~~  
"Repetition! Repetition is the mother of learning."

Sources: [GitHub](https://github.com/shivam7898/xADSM/tree/main/docs "https://github.com") | [Google Drive](https://drive.google.com/drive/folders/1KS69a-U_64OkVbIr9VKhxu97HBPzbVsz?usp=sharing "https://drive.google.com") | [OneDrive](https://1drv.ms/u/s!AjRu26ZthlQCitEmPzybZMV0tISo5w?e=JFE5ju "https://onedrive.live.com")

\textcolor{pink}{Caution:}  
The HTML File access internet to render LaTex Equations via MathJax. Thus, equations will NOT be rendered properly if the file is viewed offline. Otherwise, the file is a standalone document containing all the relevant images, tables and codes to generate them.

\textcolor{pink}{Disclaimer:}  
This document is a collection of multiple lecture notes, books, and online material. The current status of the document is 'work-in-progress' as it is incomplete and it may contain inconsistencies and errors. 

\textcolor{pink}{Attribution:}  
Proper reference to the original source may sometimes be lacking. This is regrettable and these references – if ever known to the author – will be included in later versions.

<!-- This is a comment. -->

```{r '000A', include=FALSE, cache=FALSE}
# #Knitr Setup will be run only once while building a Book or a Chapter
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
# #Packages for inclusion in Bibliography
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
# #Source ALL Functions
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
```

## Environment {.unnumbered .tabset .tabset-fade}

\textcolor{pink}{Assumption: Working directory has sub-folders named "data", "images", "code", "docs".}

### R Version {.unlisted .unnumbered}

```{r '000B-RVersion'}
# #R Version
R.version.string
```  

### Working Directory {.unlisted .unnumbered}

```{r '000C-Directory'}
# #Working Directory
getwd()
```

### Session Info {.unlisted .unnumbered}

```{r '000D-Session'}
# #Version information about R, the OS and attached or loaded packages
sessionInfo()
``` 
 
### Pandoc {.unlisted .unnumbered}

```{r '000E-Pandoc'}
# #Pandoc Version being used by RStudio
rmarkdown::pandoc_version()
```  

```{r '000F-Bib', include=FALSE}
# automatically create a bib database for R packages
knitr::write_bib(c(
  .packages(), 'bookdown', 'knitr', 'rmarkdown'
), 'packages.bib')
```

## Aside {.unlisted .unnumbered}

Originally, I created this repository because I wanted to have a single reference containing notes, codes, models, & output for data analytics. Later it was converted into a learning and collaboration tool because some of the colleagues were facing similar set of problems and the scope of this document was expanded.

The time constraint has led to issues given below. These will be corrected as and when possible.

- Tone of the document may be a little abrupt, please overlook that
- Source references are not added as much as I wanted to and there is no easy solution for this
- I have NOT explained some of the functions before their usage e.g. lapply(), identical() etc. Links will be added as and when those topics are covered
- Code has been checked only on Windows 10. For Mac or Linux, ~~if~~ when you find something that has different output or behaviour, please let me know

I am always open for suggestions. 

Thank You all for the encouragement.

Shivam Upadhyay

****
