# Data Manipulation (B11, Sep-12) {#b11}

```{r 'B11', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## Overview

- "Understanding basic data manipulation using R"
  - [Filter Rows](#row-filter-b11 "b11")
  
- Links (Ref)
  - [Transform columns as Factors](#factors-b10 "b10")
  - [Sorting Basics](#sorting-b10 "b10")
  - [Do not miss the missing values i.e. NA](#na-b10 "b10")

```{r 'B11-Flights', include=FALSE}
# #Load Data: Flights
xxflights <- f_getRDS(xxflights)
bb <- xxflights
```

## Get Help

```{r 'B11-Help', eval=FALSE}
# #To get the Help files on any Topic including 'loaded' Packages
?dplyr
?mutate
# #Help files on any Topic including functions from 'installed' but 'not loaded' Packages
?dplyr::mutate()
# #Operators need Backticks i.e. ` . In keyboards it is located below 'Esc' Key i.e. Tilde "~" Key
?`:`
# #To Get the list of All Options used by Base R (including user defined)
?options
```


## Logical Operators and Functions {#logical-b11 .tabset .tabset-fade}

- "\textcolor{pink}{|}"&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;(Or, binary, vectorized)
- "\textcolor{pink}{||}"&nbsp;&nbsp;&nbsp;&nbsp;&thinsp;(Or, binary, not vectorized)
- "\textcolor{pink}{&}"&nbsp;&nbsp;&nbsp;&nbsp;&thinsp;(And, binary, vectorized)
- "\textcolor{pink}{&&}"&nbsp;(And, binary, not vectorized)
- Functions - \textcolor{pink}{any()}, \textcolor{pink}{all()}

Overview

- Vectorised forms are "&" "|" 
  - Thus, these compare vectors elementwise and operate over complete vector length.
  - \textcolor{pink}{NA is a valid logical object}. Where a component of x or y is NA, the result will be NA if the outcome is ambiguous.
  - All components of x or y are evaluated
  - (\textcolor{pink}{recycling}) of elements occur if vector lengths are different
  - These are NOT recommended for use inside if() clauses
  - These are generally used for filtering
  - &, | do the pairwise operation in R (vs bitwise in Python, C etc.) 
- Non-vectorised forms are "&&" "||" 
  - These examine only the first element of each vector
    - \textcolor{orange}{Caution:} For these, vector length should always be 1
    - Use all() and any() to reduce the length to one
  - (\textcolor{pink}{short-circuit}) These stop execution as soon as these find at least one definite condition i.e. TRUE for ||, FALSE for &&.
    - They will not evaluate the second operand if the first operand is enough to determine the value of the expression. 
  - These are preferred in if() clauses
  - &&, || do the bitwise operation in R (vs pairwise in Python, C etc.) 
- all() and any()
  - all() : Are All Values TRUE
    - TRUE for 0-length vector
  - any() : Is at least one of the values TRUE
    - FALSE for 0-length vector
  - The value is a logical vector of length one being `TRUE`, `FALSE`, or `NA`.


### Operators {.unnumbered}

```{r 'B11-PairwiseSingle'}
NA | TRUE  #TRUE  #At least one TRUE is present
NA | FALSE #NA    #Depending upon what the unknown is, the outcome will change
NA & TRUE  #NA    #Depending upon what the unknown is, the outcome will change
NA & FALSE #FALSE #At least one FALSE is present
#
# #For length 1 vectors, output of vectorised and non-vectorised forms is same
stopifnot(all(identical(NA || TRUE, NA | TRUE), identical(NA || FALSE, NA | FALSE),
              identical(NA && TRUE, NA & TRUE), identical(NA && FALSE, NA & FALSE)))
#
# #But for vectors of >1 length, output is different
x <- 1:5
y <- 5:1
(x > 2) & (y < 3)
tryCatch( (x > 2) && (y < 3), 
          warning = function(w) {
            suppressWarnings(print((x > 2) && (y < 3)))
            cat(paste0(w))
          })
#
# # '&&' evaluates only the first element of Vector, thus caution is advised
TRUE & c(TRUE, FALSE)
TRUE & c(FALSE, FALSE)

tryCatch( TRUE && c(TRUE, FALSE), 
          warning = function(w) {
            suppressWarnings(print(TRUE && c(TRUE, FALSE)))
            cat(paste0(w))
          })
tryCatch( TRUE && c(FALSE, FALSE), 
          warning = function(w) {
            suppressWarnings(print(TRUE && c(FALSE, FALSE)))
            cat(paste0(w))
          })

TRUE && all(c(TRUE, FALSE))
TRUE && any(c(TRUE, FALSE))
```

### Evaluation {.unnumbered}

```{r 'B11-ShortCircuit'}
if(exists("x")) rm(x)
exists("x")
#
# # No short-circuit for "|" or "&", Evaluates Right and throws Error
tryCatch( TRUE | x, error = function(e) cat(paste0(e)))
tryCatch( FALSE & x, error = function(e) cat(paste0(e)))
#
# #Does not evaluate Right input because outcome already determined
tryCatch( TRUE || x, error = function(e) cat(paste0(e)))
tryCatch( FALSE && x, error = function(e) cat(paste0(e)))
# #evaluates Right input because outcome cannot be determined and throws error
tryCatch( TRUE && x, error = function(e) cat(paste0(e)))
```

### any() & all() {.unnumbered}

```{r 'B11-Any-All'}
# #any()
any(NA, TRUE)
any(NA, FALSE)
any(NA, TRUE, na.rm = TRUE)
any(NA, FALSE, na.rm = TRUE)
any(character(0))
#
# #all()
all(NA, TRUE)
all(NA, FALSE)
all(NA, TRUE, na.rm = TRUE)
all(NA, FALSE, na.rm = TRUE)
all(character(0))
```

## Relational Operators {#relational-b11}

\textcolor{pink}{$>$} , \textcolor{pink}{$<$} , \textcolor{pink}{$==$} , \textcolor{pink}{$>=$} , \textcolor{pink}{$<=$} , \textcolor{pink}{$!=$} 


## Filter  {.tabset .tabset-fade}

- \textcolor{pink}{dplyr::filter()}
  - (Masks) base::filter()
  - [Do not miss the missing values i.e. NA](#na-b10 "b10")
  - It treats comma as '&'
- \textcolor{pink}{subset()} vs. \textcolor{pink}{filter()} - 
  - \textcolor{orange}{Caution:} R Manual itself warns against usage of subset(). It is better to use \textcolor{pink}{[]} for subsetting
  - \textcolor{orange}{Caution:} NOT Verified Yet
    - subset works on matrices, however, filter does not
    - subset does not work on databases, filter does
    - subset does not drop the rownames, however, filter removes them
    - filter preserves the class of the column, subset does not
    - filter works with grouped data, subset ignores them
  - filter is stricter and thus would lead to fewer causes of unexpected outcome
- \textcolor{pink}{which()}
  - Takes a Boolean vector and returns a shorter vector containing the indices of the elements which were true.
  - If you want to know 'which' elements of a logical vector are TRUE i.e. their indices.
    - Ex: Get the position of the maximum or minimum values
  - If NA are present and you do not want them in the output
- \textcolor{pink}{with()}
  - with() is a wrapper for functions with no 'data' argument. It allows usage of function as if it had a data argument.
  
  
```{conjecture 'Object-Not-Found-01'}
\textcolor{brown}{Error in match.arg(method) : object 'day' not found}
```

- when 'dplyr' package is not loaded, base::filter() throws this error. 
- Either Load the Package (dplyr etc.) or use scope resolution '::'

### Basics {.unnumbered}

```{r 'B11-dplyrFilter'}
# #dplyr::filter() - Filter Rows based on Multiple Columns
bb_1 <- filter(bb, month == 1, day == 1)
dim(bb_1)
# #Filtering by multiple criteria within a single logical expression
stopifnot(identical(bb_1, filter(bb, month == 1 & day == 1)))
#
if(anyNA(bb_1)) {
  bb_na <- na.omit(bb_1)
  print(paste0("Note: ", length(attributes(bb_na)$na.action), " rows removed."))
} else {
  print("NA not found")
}
dim(bb_na)
```

### Conditional {.unnumbered}

```{r 'B11-ConditionalFilter'}
dim(bb)
#
# #Flights in either months of November or Decemeber
dim(bb_2 <- filter(bb, month == 11 | month == 12))
#
# #Flights with arrival delay '<= 120' or departure delay '<= 120' 
# #It excludes flights where arrival & departure BOTH are delayed by >2 hours
# #If either delay is less than 2 hours, the flight is included
dim(bb_3 <- filter(bb, arr_delay <= 120 | dep_delay <= 120))
dim(bb_4 <- filter(bb, !(arr_delay > 120 & dep_delay > 120)))
dim(bb_5 <- filter(bb, (!arr_delay > 120 | !dep_delay > 120)))
#
# #Destination to IAH or HOU
dim(bb_6 <- filter(bb, dest == "IAH" | dest == "HOU"))
dim(bb_7 <- filter(bb, dest %in% c("IAH", "HOU")))
#
# #Carrier being "UA", "US", "DL"
dim(bb_8 <- filter(bb, carrier == "UA" | carrier == "US" | carrier == "DL"))
dim(bb_9 <- filter(bb, carrier %in% c("UA", "US", "DL")))
#
# #Did not leave late (before /on time departure) but Arrived late by >2 hours
dim(bb_10 <- filter(bb, (arr_delay > 120) & !(dep_delay > 0)))
# 
# #Departed between midnight and 6 AM (inclusive)
dim(bb_11 <- filter(bb, (sched_dep_time >= 00 & sched_dep_time <= 600)))
```


### subset() {.unnumbered}

```{r 'B11-Subset'}
# #subset() - Recommendation is against its usage. Use either '[]' or filter()
dim(bb_12 <- subset(bb, month == 1 | !(dep_delay >= 120), 
                    select = c("flight", "arr_delay")))
dim(bb_13 <- subset(bb, month == 1 | !(dep_delay >= 120) | carrier == "DL", 
                select = c("flight", "arr_delay")))
```


### Filter Rows {#row-filter-b11 .unnumbered}

```{r 'B11-SubsetRow'}
# #Data: mtcars, 32x11, "mpg, cyl, disp, hp, drat, wt, qsec, vs, am, gear, carb"
bb <- aa <- mtcars
#str(bb)
#summary(bb)
# #
# #Avoid subset()
ii <- subset(bb, wt > 2 & wt < 3)
# #which() 
jj <- bb[which(bb$wt > 2 & bb$wt <= 3), ]
#
# #which() select only TRUE and NOT the NA
(1:2)[which(c(TRUE, NA))]
(1:2)[c(TRUE, NA)]
#
# #which() is faster than head() 
ee <- bb[which(bb$wt > 2 & bb$wt <= 3)[1:6], ] 
ff <- head(bb[bb$wt > 2 & bb$wt <= 3, ], 6)
stopifnot(identical(ee, ff))
#
# #Normal Filter using [] operator
kk <- bb[bb$wt > 2 & bb$wt <= 3, ]
#
# #with()
ll <- with(bb, bb[wt > 2 & wt <= 3, ])
#
# #filter()
mm <- bb %>% filter(wt > 2 & wt <= 3)
#
stopifnot(all(identical(ii, jj), identical(ii, kk), identical(ii, ll), identical(ii, mm)))
#
# #Another set of equivalent operations for OR 
ii <- subset(bb, cyl == 4 | cyl == 6)
jj <- bb[bb$cyl %in% c(4, 6), ]
kk <- bb[which(bb$cyl %in% c(4, 6)), ]
ll <- bb %>% filter(cyl == 4 | cyl == 6)
mm <- bb %>% filter(cyl %in% c(4, 6))
#
stopifnot(all(identical(ii, jj), identical(ii, kk), identical(ii, ll), identical(ii, mm)))
#
# #General Conditional Subsetting on Flights data
bb <- xxflights
dim(bb)
#
dim(bb[which(bb$day == 1 & !(bb$month ==1)), ])
dim(bb[which(bb$day == 1 | bb$month ==1), ])
dim(bb[which(bb$day == 1 & bb$month ==1), ])
dim(bb[which(bb$day == 1, bb$month ==1), ])
dim(bb[which(bb$day == 1 & !(bb$carrier == "DL")), ])
dim(bb[which(bb$day == 1 | bb$carrier == "DL"), ])
dim(bb[which(bb$day == 1 & bb$carrier == "DL"), ])
dim(bb[which(bb$day == 1, bb$carrier == "DL"), ])
```



## Subsetting {.tabset .tabset-fade}

\textcolor{pink}{$[ \ \ ]$} , \textcolor{pink}{$[[ \ \ ]]$} , \textcolor{pink}{$\$$}

- Extract or Replace Parts of an Object
  - Operators acting on vectors, matrices, arrays and lists to extract or replace parts. 
  - The most important distinction between "[ ]", "[[ ]]" and "$" is that the "[ ]" can select more than one element whereas the other two select a single element. 
  - "$" does not allow computed indices, whereas "[[ ]]" does.
  - Subsetting (except by an empty index) will drop all attributes except names, dim and dimnames. Indexing will keep them.
  
```{conjecture 'Comparison-possible'}
\textcolor{brown}{Error in day == 1 : comparison (1) is possible only for atomic and list types}
```

- It occurs when the data is not available i.e. column name is NOT found
- It might happen when the original code assumed that the dataframe is attached
- Either attach the dataframe (NOT Recommended) or use "$" to access column names

\textcolor{pink}{dplyr::select()}

- It can use Range "\textcolor{pink}{:}", Not "\textcolor{pink}{!}", And "\textcolor{pink}{&}", Or "\textcolor{pink}{|}" 
- Selection Helpers
  - everything(): Matches all variables.
  - last_col(): Select last variable, possibly with an offset.
- These helpers select variables by matching patterns in their names:
  - starts_with(): Starts with a prefix.
  - ends_with(): Ends with a suffix.
  - contains(): Contains a literal string.
  - matches(): Matches a regular expression.
  - num_range(): Matches a numerical range like x01, x02, x03.
- These helpers select variables from a character vector:
  - all_of(): Matches variable names in a character vector. All names must be present, otherwise an out-of-bounds error is thrown.
  - any_of(): Same as all_of(), except that no error is thrown for names that do not exist.
- This helper selects variables with a function:
  - where(): Applies a function to all variables and selects those for which the function returns TRUE.


### Cols {.unnumbered}

```{r 'B11-SubsetCol'}
dim(bb)
#
# #Subset Consecutive Columns using Colon
stopifnot(identical(bb[ , 2:5], bb[ , -c(1, 6:ncol(bb))]))
#
# #dplyr::select()
bb_14 <- select(bb, year:day, arr_delay, dep_delay, distance, air_time)
bb_15 <- bb %>% select(year:day, arr_delay, dep_delay, distance, air_time)
stopifnot(identical(bb_14, bb_15))
```


## Grouped Summary {.tabset .tabset-fade}

- \textcolor{pink}{dplyr::summarise()} or dplyr::summarize()
- \textcolor{pink}{dplyr::group_by()}
  - It converts an existing Tibble into a grouped Tibble where operations are performed "by group". 
  - \textcolor{pink}{ungroup()} removes grouping.
  - \textcolor{pink}{n()} gives the number of observations in the current group.

### Summarise {.unnumbered}

```{r 'B11-Summarise'}
bb <- xxflights
# #dplyr::summarise() & dplyr::summarize() are same
# #Get the mean of a column with NA excluded
#
summarize(bb, delay_mean = mean(dep_delay, na.rm = TRUE))
#
# #base::summary()
summary(bb$dep_delay)
#
# #Grouped Summary
by_ymd <- group_by(bb, year, month, day)
mysum <- summarize(by_ymd, 
                   dep_delay_mean = mean(dep_delay, na.rm = TRUE), 
                   arr_delay_mean = mean(arr_delay, na.rm = TRUE),
                   .groups = "keep")
# #Equivalent 
bb %>% 
  group_by(year, month, day) %>% 
  summarize(dep_delay_mean = mean(dep_delay, na.rm = TRUE), 
			arr_delay_mean = mean(arr_delay, na.rm = TRUE),
			.groups= "keep")
```


### group_by() {.unnumbered}

```{r 'B11-SummaryDistance'}
# #Get delay grouped by distance 'Distance between airports, in miles.'
summary(bb$distance)
#
# #How many unique values are present in this numeric data i.e. factors
str(as.factor(bb$distance))
str(sort(unique(bb$distance)))
bb %>% 
  group_by(distance) %>% 
  summarize(count = n(),
            dep_delay_mean = mean(dep_delay, na.rm = TRUE), 
            arr_delay_mean = mean(arr_delay, na.rm = TRUE),
            .groups= "keep")
#
# #For distance =17, there is only 1 flight and that too has NA, so the mean is NaN
bb[bb$distance == 17, ]
#
# #In general, Flight to any destination (ex: ABQ) has travelled same distance (1826)
unique(bb %>% filter(dest == "ABQ") %>% select(distance))
#
# #Mean Delays for Destinations with more than 1000 miles distance
bb %>% 
  group_by(dest) %>% 
  filter(distance > 1000) %>% 
  summarize(count = n(), 
            distance_mean = mean(distance, na.rm = TRUE),
            dep_delay_mean = mean(dep_delay, na.rm = TRUE), 
            arr_delay_mean = mean(arr_delay, na.rm = TRUE))
```


## Mutate {.tabset .tabset-fade}

- \textcolor{pink}{dplyr::mutate()}
  - Newly created variables are available immediately
  - New variables overwrite existing variables of the same name. 
  - Variables can be removed by setting their value to NULL.
  - mutate() adds new variables and preserves existing ones
    - mutate() can also keep or drop column according to the .keep argument.
  - transmute() adds new variables and drops existing ones. 

```{conjecture 'UseMethod-No-applicable-method'}
\textcolor{brown}{Error in UseMethod("select") : no applicable method for 'select' applied to an object of class "function"}
```

- Run 'str(MyObject)' to check if 'MyObject' exists, looks as expected and R is not finding something else. 
- Most probably R reserved keyword 'data' was called in place of the actual 'data'.
- To minimise this type of Error, do not use the keywords which match with Base R Functions e.g. 'data' (Function in utils) or 'df' (function in stats)

```{conjecture 'Object-Not-Found-02'}
\textcolor{brown}{Error: Problem with mutate() column ... column object ... not found}
```

- Run 'str(MyObject)' to check if the column exists in the dataset 
- \textcolor{orange}{Caution:} if the dataset was attached earlier, then R will NOT throw this error. However, later when the code is being executed in a clean environment, it will fail. To avoid this, it is recommended to use proper scope resolution and to avoid attaching the dataset (if possible)

```{r 'B11-Mutate'}
dim(bb)
#
bb_16 <- select(bb, year:day, arr_delay, dep_delay, distance, air_time)
bb_17 <- mutate(bb_16,
       gain = arr_delay - dep_delay,
       speed = distance / air_time * 60,
       hours = air_time / 60,
       gain_per_hour = gain / hours)
# #Equivalent
bb %>% 
  select(year:day, arr_delay, dep_delay, distance, air_time) %>% 
  mutate(gain = arr_delay - dep_delay,
         speed = distance / air_time * 60,
         hours = air_time / 60,
         gain_per_hour = gain / hours)
```


## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'B11-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll, bb_1, bb_10, bb_11, bb_12, bb_13, bb_14, bb_15, bb_16, bb_17, bb_2, bb_3, bb_4, bb_5, bb_6, bb_7, bb_8, bb_9, bb_na, by_ymd, mysum, x, xxflights, y, ff)
```

```{r 'B11-Validation', include=FALSE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
