# Quiz (B15, Oct-17) {#b15}

```{r 'B15', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## Short Quiz

1. In hypothesis testing,
    a. the smaller the Type I error, the smaller the Type II error will be
    a. the smaller the Type I error, the larger the Type II error will be
    a. Type II error will not be affected by Type I error
    a. the sum of Type I and Type II errors must equal to 1
	  - Answer: \textcolor{black}{b}
1. What type of error occurs if you accept ${H_0}$ when, in fact, it is not true
    a. Type II
    a. Type I
    a. either Type I or Type II, depending on the level of significance
    a. either Type I or Type II, depending on whether the test is one tail or two tail
	  - Answer: \textcolor{black}{a}
1. If the level of significance of a hypothesis test is raised from .01 to .05, the probability of a Type II error
    a. will also increase from .01 to .05
    a. will not change
    a. will decrease
    a. will increase
	  - Answer: \textcolor{black}{c}
1. The sum of the values of ${\alpha}$ and ${\beta}$
    a. always add up to 1.0
    a. always add up to 0.5
    a. is the probability of Type II error
    a. None of these alternatives is correct
	  - Answer: \textcolor{black}{d}
1. Following the p-value approach, the null hypothesis is rejected if 
    a. p-value less than or equal to ${\alpha}$
    a. ${\alpha}$ < p-value
    a. p-value > ${\alpha}$
    a. p-value = 1 - ${\alpha}$
	  - Answer: \textcolor{black}{a}
1. The average manufacturing work week in metropolitan Chattanooga was 40.1 hours last year. It is believed that the recession has led to a reduction in the average work week. To test the validity of this belief, which of the following stated hypothesis formulation is true.
    a. ${H_0} : {\mu} < 40.1 \quad {H_a} : {\mu} \geq 40.1$
    a. ${H_0} : {\mu} \geq 40.1 \quad {H_a} : {\mu} < 40.1$
    a. ${H_0} : {\mu} > 40.1 \quad {H_a} : {\mu} \leq 40.1$
    a. ${H_0} : {\mu} = 40.1 \quad {H_a} : {\mu} \neq 40.1$
	  - Answer: \textcolor{black}{b}
1. Statement about a population developed for the purpose of testing is called:
    a. Hypothesis
    a. Hypothesis testing
    a. Level of significance
    a. Test-statistic
	  - Answer: \textcolor{black}{a}
1. The probability of rejecting the null hypothesis when it is true is called: 
    a. Level of confidence
    a. Level of significance
    a. Power of the test
    a. Difficult to tell
	  - Answer: \textcolor{black}{b}
1. The dividing point between the region where the null hypothesis is rejected and the region where it is not rejected is said to be:
    a. Critical region
    a. Critical value
    a. Acceptance region
    a. Significant region
	  - Answer: \textcolor{black}{b}
1. If the critical region is located equally in both sides of the sampling distribution of test-statistic, the test is called:
    a. One tailed
    a. Two tailed
    a. Right tailed
    a. Left tailed
	  - Answer: \textcolor{black}{b}
1. Test of hypothesis ${H_0}$: ${\mu} \leq 50$ against ${H_a}$: ${\mu} > 50$ leads to:
    a. Left-tailed test
    a. Right-tailed test
    a. Two-tailed test
    a. Difficult to tell
	  - Answer: \textcolor{black}{b}
1. Test of hypothesis ${H_0}$: ${\mu} = 20$ against ${H_1}$: ${\mu} < 20$ leads to:
    a. Right one-sided test
    a. Left one-sided test
    a. Two-sided test
    a. All of the above
	  - Answer: \textcolor{black}{b}
1. A failed student has been promoted by an examiner; it is an example of: 
    a. Type-I error
    a. Type-II error
    a. Unbiased decision
    a. Difficult to tell
	  - Answer: \textcolor{black}{b}
1. The probability of accepting ${H_0}$ when it is True is called: 
    a. Power of the test
    a. Size of the test
    a. Level of confidence
    a. Confidence coefficient
	  - Answer: \textcolor{black}{d}
1. Power of a test is directly related to:
    a. Type-I error
    a. Type-II error
    a. Both (a) and (b)
    a. Neither (a) nor (b)
	  - Answer: \textcolor{black}{b}
  
## Case: 



## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'B15-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ee, hh, ii, jj, kk, ll, mm, nn, oo, rr, vv, xx, yy, zz)
```

```{r 'B15-Validation', include=TRUE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
