# Regresssion (B25, Dec-26) {#b25}

```{r 'B25', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## Overview

- "Supervised Learning Algorithm: Regression"


## Linear Regression

```{r 'B25D01', comment="", echo=FALSE, results='asis'}
f_getDef("Simple-Linear-Regression") #dddd
```

- ${Y}$: Scalar response, Dependent variable, Outcome variable, Target, Predicted
- ${X}$: Explanatory variables, Independent variables, Antecendent variables, Predictors

- It is applied when the objective is to predict the outcome variable based on the antecendent variables
- Predicted (Y) should be continuous, but Predictors (X) can be either continuous or categorical 
  - Ex: Salary (Y) is a function of Age, Gender, Education, Years of Experience
  - Ex: Consumption (Y) is a function of Income (X)
  - We are interested in $Y = mX + C$ where ${m}$ is the slope of the line and C is the y-intercept
    - Slope: What is the change in Y for unit change in X
  - Because there will be some error ${\epsilon}$, thus the equation is given by $y = {\alpha} + {\beta}x + {\epsilon}$ where ${\alpha}$ is the average Y when X are zero.
  - NOTE Equation can also be given by ${y} = {\beta}_0 + {\beta}_1 {x} + {\epsilon}$. In that case assume ${\alpha} = {\beta}_0$ and ${\beta} = {\beta}_1$.

- \textcolor{pink}{Question:} ${\alpha}$ is constant for a given set of data
  - Yes

- Refer [Why Slope is more important than the intercept](#beta1-c14 "c14")
- Simple Linear Regression is also known as \textcolor{pink}{Bivariate Regression}. i.e. Single Y, Single X
- When there are Single Y and Multiple X, it is called \textcolor{pink}{Multiple Linear Regression}
  - Equation: ${y} = {\beta}_0 + {\beta}_1 {x}_1 + {\beta}_2 {x}_2 + \ldots + {\epsilon}$


## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'B25-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll)
```

```{r 'B25-Validation', include=TRUE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
