# Linear Regression (B27, Jan-09) {#b27}

```{r 'B27', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## Overview

- "Machine learning using linear regression"
  - Skipped "15:48-15:56" because interactive function 'file.choose()' or Import is difficult to show in Notebook.

## Data: Hospital {.tabset .tabset-fade}

\textcolor{pink}{Please import the "B27-Hospital.xlsx"}

```{r 'B27-namesHospital', include=FALSE}
# #Object Names for each sheet
namesXL <- c("xxB27Hdesc", "xxB27Hraw", "xxB27Hmod")
```

```{r 'B27-getHospital', include=FALSE, eval=FALSE}
# #Path of Object, FileName and MD5Sum
#tools::md5sum(paste0(.z$XL, "B27-Hospital.xlsx"))
xxB27Hospital <- f_getObject("xxB27Hospital", "B27-Hospital.xlsx", "176788c58511ea4e98739a5314a55aab")
```

```{r 'B27-XLobjects', include=FALSE, eval=FALSE}
# #Create Separate Tibbles for Each Sheet
# #Separate Objects, excluding the First Sheet
for(ii in seq_along(namesXL)){
  assign(namesXL[ii], xxB27Hospital[[ii + 1]])
}
# #Save Binary Files
for(ii in seq_along(namesXL)){
  saveRDS(eval(parse(text = namesXL[ii])), paste0(.z$XL, namesXL[ii], ".rds"))
}
```

```{r 'B27-GetHospital', include=FALSE}
for(ii in seq_along(namesXL)){
  assign(namesXL[ii], readRDS(paste0(.z$XL, namesXL[ii], ".rds")))
}
```

```{r 'B27-List', include=FALSE, eval=FALSE}
# #Dimensions of these datasets
str(lapply(namesXL, function(x) {dim(eval(parse(text = x)))}))
```

### EDA {.unlisted .unnumbered}

- About: [248, 24]
  - Sanitise Column Names for ease of use
  - NA in 7 columns 
  - ~~If implant was not used then implant cost was changed from 0 to NA~~
    - Regression really does not like NA.
  - Dropped "State at the Time of Arrival" because only 1/248 has "Confused", all others are "Alert"
  - Converted Low BP from Character to Numeric
    - Low BP and High BP have 23 NA for same patients
  - Converted Binary Variables to dummy variables i.e. iFemale, iMarried, iEmergency, iImplant
  - History has Typo in case so convert all to lower, change labels and made "ht1" as reference because it is the most frequent
    - NA in History were merged with already available "other" category


```{r 'B27-PrepHospital'}
# #Sanitise Headers
aa <- xxB27Hraw %>% 
  rename_with(make.names) %>% rename_with(~ tolower(gsub(".", "_", .x, fixed = TRUE))) 
names(aa) <- c("ID", "Age", "iFemale", "iMarried", "Complaints", "Weight", "Height", "Pulse", 
               "HighBP", "LowBP", "Respiratory", "History", "Hemoglobin", "Urea", "Creatinine", 
               "Arrival", "State", "iEmergency", "TC", "Stay", "ICU", "Ward", "iImplant", "ImpCost")
#
# #Drop | To Numeric | Relocate | Flags | Lower Case | Replace NA | 
# #Levels & Labels | Replace String | To Factor
# #To Remove Multicollinearity: Dropped Arrival, Dropped iImplant, Drop Stay (~ ICU + Ward)
# #Reverted ImpCost to original (0) for No implant 
# #Changed NA of Complaints to 'other'
# #Remove Rows containing NA in HighBP, LowBP, Hemoglobin, Urea, Creatinine
#
bb <- aa %>% 
  select(-State) %>% 
  mutate(across(LowBP, as.numeric)) %>% 
  relocate(TC, .after = ID) %>% 
  mutate(iFemale = ifelse(iFemale == "F", 1, 0)) %>% 
  mutate(iMarried = ifelse(iMarried == "MARRIED", 1, 0)) %>% 
  mutate(iEmergency = ifelse(iEmergency == "EMERGENCY", 1, 0)) %>% 
  mutate(iImplant = ifelse(iImplant == "Y", 1, 0)) %>% 
  mutate(across(Complaints, coalesce, "other")) %>% 
  mutate(across(History, tolower)) %>% 
  mutate(across(History, coalesce, "other")) %>% 
  mutate(across(History, factor, 
levels = c("hypertension1", "hypertension2", "hypertension3", "diabetes1", "diabetes2", "other"), 
labels = c("ht1", "ht2", "ht3", "db1", "db2", "other"))) %>% 
  mutate(Arrival = str_replace_all(Arrival, c(" " = "_"))) %>% 
  mutate(Arrival = factor(Arrival, levels = c("WALKED_IN", "AMBULANCE", "TRANSFERRED"), 
                          labels = c("Walked", "Ambulance", "Transferred"))) %>% 
  mutate(Complaints = str_replace_all(Complaints, 
                        c(" " = "_", "-" = "_", "__" = "_", "other_" = "o_"))) %>% 
  mutate(across(Complaints, factor)) %>% 
  mutate(across(Complaints, relevel, ref = "o_heart")) %>% 
  #select(-c(Arrival, iImplant, Stay)) %>% 
  drop_na(HighBP, LowBP, Hemoglobin, Urea, Creatinine)
#
xsyw <- bb
```

### Structure {.unlisted .unnumbered}

```{r 'B27-Temps', include=FALSE, eval=FALSE}
# #Count NA 
if(FALSE) colSums(is.na(bb)) %>% as_tibble(rownames = "Cols") %>% filter(value > 0)
#
if(FALSE) bb %>% select(where(is.numeric)) %>% summary()
if(FALSE) bb %>% select(isFemale) %>% slice(1:10)
if(FALSE) cat('"', paste0(levels(bb$Arrival), collapse = '", "'), '"\n', sep = '')
if(FALSE) bb %>% count(Complaints)
if(FALSE) bb %>% filter(is.na(bpl)) %>% select(id, bph, bpl)
if(FALSE) bb %>% select(isImplant, imp_c) %>% summary()
if(FALSE) bb %>% select(iFemale, iMarried, iEmergency, iImplant) %>% 
  pivot_longer(everything()) %>% 
  count(name, value) %>% 
  pivot_wider(names_from = value, values_from = n) 
```

```{r 'B27-strH'}
str(bb)
```

### Summary {.unlisted .unnumbered}

```{r 'B27-summaryH'}
summary(bb)
```

## Dummies {.tabset .tabset-fade}

- Refer [Omitting the rows with missing data is not recommended](#imputation-b16 "b16")

- Following is applicable to First Run of the Model which had Multicollinearity issues and thus was modified further.
  - Note: The dummy object created here contains 40 columns. It is slightly different from what professor has created with 41 columns.
  - I have skipped 'state on arrival' and kept 'ID' for now
  - I have merged NA in 'history' column to 'other' category
  - Further, the Reference Level is different in some dummy variables
  - I have NOT removed rows containing NA. Those will be removed just before the Model building 
    - Ideally these should be imputed but that is "ForLater"
      - Medical information is difficult to impute
    - Number of Rows with NA is higher (211) in my case because I have replaced 0 in 'Implant Cost' to NA
  - If the original categorical column has 'NA' the dummy creation lead to extra 'NA' column. 
    - The other columns would have {0, 1, NA} in this case. For now those NA are being set to 0 manually. Impact of keeping these NA as it is and handling it inside model would be observed "ForLater"

### Create {.unlisted .unnumbered}

```{r 'B27-Dummy'}
# #Create Dummies | Replaced All NA in original data so no replacement in dummy columns |
dum_xsyw <- dummy_cols(xsyw, remove_first_dummy = TRUE, remove_selected_columns = TRUE) 
```

### Structure {.unlisted .unnumbered}

```{r 'B27-StrHd'}
names(dum_xsyw)
str(dum_xsyw)
```

### NA {.unlisted .unnumbered}

```{r 'B27-NA'}
sum(!complete.cases(dum_xsyw))
colSums(is.na(dum_xsyw)) %>% as_tibble(rownames = "Cols") %>% filter(value > 0)
```

## Model {.tabset .tabset-fade}

- First Run of Model with 40 variables
  - Multicollinear

```{conjecture 'vif-aliased'}
\textcolor{brown}{Error in vif.default(...) : there are aliased coefficients in the model}
```

- There is Perfect Multicollinearity issue in the Model Variables
- To identify which variables are the culprits, run use \textcolor{pink}{alias()}
  - It tells that 'Arrival_Ambulance' has perfect collinearity with 'iEmergency'
    - ~~In this case we should drop the Arrival Variable completely~~
    - It turns out that the NA were the main problem. Multicollinearity can be handled by step()
  - Ultimately all NA were removed from Continuous Variables and NA in Complaints were converted to Other

```{conjecture 'step-rows-changed'}
\textcolor{brown}{Error in step(...) : number of rows in use has changed: remove missing values}
```

- If the original model was created with using 'na.action' of lm(), step() throws this error
  - Remove all NA from original dataset

### Model 4 {.unlisted .unnumbered}

```{r 'B27-ModelH'}
# #Partition Data
set.seed(3)
idx_xsyw <- sample.int(n = nrow(dum_xsyw), size = floor(0.8 * nrow(dum_xsyw)), replace = FALSE)
#
# #Drop the ID Column from Training
train_xsyw <- dum_xsyw[idx_xsyw, ] %>% select(-ID) 
test_xsyw  <- dum_xsyw[-idx_xsyw, ]
#
mod_xsyw <- lm(TC ~ ., data = train_xsyw)
if(FALSE) f_pNum(summary(mod_xsyw)$coefficients) %>% as_tibble(rownames = "DummyParVsRef") %>% 
  rename(pVal = "Pr(>|t|)") %>% 
  mutate(pVal = ifelse(pVal < 0.001, 0, pVal), isSig = ifelse(pVal < 0.05, TRUE, FALSE))
#
stp_xsyw <- step(mod_xsyw, direction = "backward", trace = 0)
if(TRUE) f_pNum(summary(stp_xsyw)$coefficients) %>% as_tibble(rownames = "DummyParVsRef") %>% 
  rename(pVal = "Pr(>|t|)") %>% 
  mutate(pVal = ifelse(pVal < 0.001, 0, pVal), isSig = ifelse(pVal < 0.05, TRUE, FALSE))
#
summary(stp_xsyw)$adj.r.squared
names(stp_xsyw$coefficients)
```

### VIF {.unlisted .unnumbered}

```{r 'B27-VifH'}
ii <- vif(mod_xsyw) 
ii[ii > 5]
#
ii <- vif(stp_xsyw) 
ii[ii > 5]
```

### Multicollinear {.unlisted .unnumbered}

```{r 'B27-Model2Coll', eval=FALSE}
# #Had to Remove all NA from Original Dataset to handle the Error
mm <- alias(mod_xsyw)$Complete
mm[rowSums(mm^2) != 0, colSums(mm^2) != 0]
#
if(FALSE) xsyw %>% filter(History == "db1") %>% select(ID, History, Complaints, iEmergency)
```

## Model 5

- To Explore: What happens if we remove the NA of "Complaints" Column (earlier those were converted to 'Other')
  - Different Variables become significant 
    - However, it might be due to change in Train Data because number of Rows has changed. The included and excluded rows have changed even if we are re-setting the seed immediately before the sampling.
    - "ForLater" Take a dataset do the same exercise but filter on Train data (not the original dataset as done here). It would show whether the variables included remain same or not.

- Comparison of Models
  - ~~Model 2 (Excluding Arrival, iImplant, Stay) & (Complaints NA to Other)~~ 
    - $R_a^2 = 0.873$
    - No VIF > 5
    - Variables: 10
      - (Intercept), Pulse, Hemoglobin, Urea, Creatinine, ICU, Ward, ImpCost, Complaints_CAD_TVD, Complaints_o_general, History_db2
      - Common: Pulse, ICU, Ward, ImpCost, Complaints_o_general, History_db2
      - NOT in Model 3: Hemoglobin, Urea, Creatinine, Complaints_CAD_TVD
  - ~~Model 3 (Excluding Arrival, iImplant, Stay) & (Excluding NA of Complaints)~~ 
    - $R_a^2 = 0.863$
    - No VIF > 5
    - Variables: 9
      - (Intercept), Age, Height, Pulse, ICU, Ward, ImpCost, Complaints_o_general, Complaints_PM_VSD, History_db2
      - Common: Pulse, ICU, Ward, ImpCost, Complaints_o_general, History_db2
      - NOT in Model 2: Age, Height, Complaints_PM_VSD
  - Model 4 (Complaints NA to Other)
    - $R_a^2 = 0.873$
    - No VIF > 5
    - Variables: 10 (Only Stay replaced Ward compared to Model 2)
      - (Intercept), Pulse, Hemoglobin, Urea, Creatinine, Stay, ICU, ImpCost, Complaints_CAD_TVD, Complaints_o_general, History_db2
      - Common: Pulse, ICU, Stay, ImpCost, Complaints_o_general, History_db2
      - NOT in Model 5: Hemoglobin, Urea, Creatinine, Complaints_CAD_TVD
  - Model 5 (Excluding NA of Complaints) 
    - $R_a^2 = 0.861$
    - No VIF > 5
    - Variables: 8 (Stay replaced Ward, Tertology replaced VSD and Dropped Height compared to Model 3)
      - (Intercept), Age, Stay, ICU, ImpCost, Complaints_o_general, Complaints_o_tertalogy, History_db2
      - Common: Pulse, ICU, Stay, ImpCost, Complaints_o_general, History_db2
      - NOT in Model 4: Age, Height, Complaints_o_tertalogy


```{r 'B27-Model3'}
# #Remove the Original NA (that were converted to other) in Complaints
ii <- xsyw %>% filter(Complaints != "other") %>% mutate(across(Complaints, droplevels))
dum_ii <- dummy_cols(ii, remove_first_dummy = TRUE, remove_selected_columns = TRUE) 
# #Partition Data
set.seed(3)
idx_ii <- sample.int(n = nrow(dum_ii), size = floor(0.8 * nrow(dum_ii)), replace = FALSE)
#
# #Drop the ID Column from Training
train_ii <- dum_ii[idx_ii, ] %>% select(-ID) 
test_ii  <- dum_ii[-idx_ii, ]
#
mod_ii <- lm(TC ~ ., data = train_ii)
if(FALSE) f_pNum(summary(mod_ii)$coefficients) %>% as_tibble(rownames = "DummyParVsRef") %>% 
  rename(pVal = "Pr(>|t|)") %>% 
  mutate(pVal = ifelse(pVal < 0.001, 0, pVal), isSig = ifelse(pVal < 0.05, TRUE, FALSE))
#
stp_ii <- step(mod_ii, direction = "backward", trace = 0)
if(FALSE) f_pNum(summary(stp_ii)$coefficients) %>% as_tibble(rownames = "DummyParVsRef") %>% 
  rename(pVal = "Pr(>|t|)") %>% 
  mutate(pVal = ifelse(pVal < 0.001, 0, pVal), isSig = ifelse(pVal < 0.05, TRUE, FALSE))
#
vif_ii <- vif(stp_ii)
vif_ii[vif_ii > 5]
#
summary(stp_ii)$adj.r.squared
names(stp_ii$coefficients)
```


## Normality 

```{r 'B27D01', comment="", echo=FALSE, results='asis'}
f_getDef("Simple-Linear-Regression-Assumption-Summary") #dddd
```

- To check the normality assumption of each $x_i$, we can look at the normality of error term $\epsilon$
  - The $\epsilon$ will vary depending upon the $\beta_i$ and the $\beta_i$ vary depending upon the $x_i$
  - So if $x_i$ follows an assumption, then $\epsilon$ should follow that assumption
  - If we presume that $x_i$ is follow Normal Distribution then it implies that $\epsilon$ will also follow Normal Distribution i.e. $x_i \in \mathcal{N} : \epsilon \in \mathcal{N}$
  - Explanation
    - $\beta_i$ should explain maximum variability, so there is an association between $\beta_i$ and $\epsilon$. Change in $\beta_i$ would impact $\epsilon$.
    - $\epsilon$ is the part unexplained by $x_i$. So, it gets affected by $x_i$
  - If $\epsilon$ is not distributed normally then at least one of the $x_i$ is not distributed normally and we would need to take corrective action


- \textcolor{pink}{Question:} What are the causes of non-normality in error term $\epsilon$
  - One of the common cause is Outliers

- \textcolor{pink}{Question:} "ForLater" This will tell us about only those $x_i$ which were included in the model. It would not comment on the variables that got dropped during the model or step model creation


```{r 'B27-Normal'}
# #Normality Test
test_shapiro <- shapiro.test(mod_xsyw$residuals)
test_shapiro
#
if(test_shapiro$p.value > 0.05) cat("Normal. H0.\n") else cat("Not Normal. Ha.\n")
```

## Semi-Log Transformation

- To decrease the skewness, we need to transform Y.
  - Semi-log Transformation: Take Log only on Y
  - Log-log Transformation: If situation does not improve, we will take log of Y and all $x_i$

- \textcolor{pink}{Question:} Can we do the Normalisation
  - No, Scaling or Normalisation does not reduce the skewness.

- \textcolor{pink}{Question:} "ForLater" During log-log how would we take log of dummy variables {0, 1}
  - (Aside) Probably, we do not apply log on dummy variables. As we have seen earlier, their variance is already controlled.

- Semi-log Transformation 
  - Model still is NOT following Normality
  - $R_a^2 = 0.828$
    - VIF > 5 is for Two Variables: Age, iMarried (It can be ignored)
    - Variables: 18 (8 more than original Model 4)

- \textcolor{pink}{Question:} "ForLater" Now both iImplant and ImpCost are present in the Model without any problem. Even though iImplant = 0 (Flag) directly corresponds with ImpCost = 0 (Continuous). This might be the correct way to do regression after all!

```{r 'B27-TransformSemi'}
# #Run Log Transformed Y
mod_xsyw_log <- lm(log(TC) ~ ., data = train_xsyw)
stp_xsyw_log <- step(mod_xsyw_log, direction = "backward", trace = 0)
if(FALSE) f_pNum(summary(stp_xsyw_log)$coefficients) %>% as_tibble(rownames = "DummyParVsRef") %>% 
  rename(pVal = "Pr(>|t|)") %>% 
  mutate(pVal = ifelse(pVal < 0.001, 0, pVal), isSig = ifelse(pVal < 0.05, TRUE, FALSE))
#
vif_ii <- vif(stp_xsyw_log)
vif_ii[vif_ii > 5]
#
summary(stp_xsyw_log)$adj.r.squared
names(stp_xsyw_log$coefficients)
#
# #Normality
test_shapiro <- shapiro.test(stp_xsyw_log$residuals)
test_shapiro
#
if(test_shapiro$p.value > 0.05) cat("Normal. H0.\n") else cat("Not Normal. Ha.\n")
```

## Log-Log Transformation - Normal

- \textcolor{pink}{Question:} Sometimes step() does not remove some variable which are not significant (as given by p-value)
  - Presence of these variables help other variables in making significant contribution to the model

- Log-log Transformation 
  - \textcolor{pink}{Model is NORMAL.}
    - Added +1 to 4 variables with 0 values to handle 'log(0) = -Inf' problem. Two of them anyway were dropped due to multicollinearity.
  - $R_a^2 = 0.814$
    - No VIF > 5
    - Variables: 16 

```{r 'B27-TransformLogLog'}
# #Run Log Transformed Y and X
train_log <- train_xsyw %>% 
  mutate(across(c(Stay, ICU, Ward, ImpCost), ~ . + 1)) %>% 
  mutate(across(c(TC, Age, Weight, Height, Pulse, HighBP, LowBP, Respiratory, Hemoglobin, Urea,
                  Creatinine, Stay, ICU, Ward, ImpCost), log)) %>% 
  select(-c(ImpCost, Stay))
#
mod_xsyw_lll <- lm(TC ~ ., data = train_log)
stp_xsyw_lll <- step(mod_xsyw_lll, direction = "backward", trace = 0)
if(FALSE) f_pNum(summary(stp_xsyw_lll)$coefficients) %>% as_tibble(rownames = "DummyParVsRef") %>% 
  rename(pVal = "Pr(>|t|)") %>% 
  mutate(pVal = ifelse(pVal < 0.001, 0, pVal), isSig = ifelse(pVal < 0.05, TRUE, FALSE))
#
vif_ii <- vif(stp_xsyw_lll)
vif_ii[vif_ii > 5]
#
summary(stp_xsyw_lll)$adj.r.squared
names(stp_xsyw_lll$coefficients)
#
# #Normality
test_shapiro <- shapiro.test(stp_xsyw_lll$residuals)
test_shapiro
#
if(test_shapiro$p.value > 0.05) cat("Normal. H0.\n") else cat("Not Normal. Ha.\n")
```


## Predict

- Subtract Y from 'exp(predicted)' to get the RMSE etc. in case of log transformed Y.

```{r 'B27-Predict'}
# #Apply Same Transformation on Test Data (except the Actual Y)
test_log <- test_xsyw %>% 
  mutate(across(c(Stay, ICU, Ward, ImpCost), ~ . + 1)) %>% 
  mutate(across(c(#TC, 
                  Age, Weight, Height, Pulse, HighBP, LowBP, Respiratory, Hemoglobin, Urea,
                  Creatinine, Stay, ICU, Ward, ImpCost), log)) %>% 
  select(-c(ImpCost, Stay))
#
res_lll <- test_log %>% 
  mutate(CalYlog = predict(stp_xsyw_lll, .), Y_Yc = TC - exp(CalYlog))
#
summary(res_lll$Y_Yc)
#
# #RMSE: Root Mean Squared Error
#res_w %>% summarise(across(everything(), ~ sqrt(mean((.)^2))))
sqrt(mean((res_lll$Y_Yc)^2))

#
# #MAE: Mean Absolute Error (MAE)
#res_w %>% summarise(across(everything(), ~ mean(abs(.))))
mean(abs(res_lll$Y_Yc))
```


## Glance

```{r 'B27-Glance'}
lapply(list(stp_xsyw_lll, stp_xsyw_log, mod_xsyw, stp_ii), glance)
```

## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'B27-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll, ff, namesXL, dum_ii, dum_xsyw, idx_ii, idx_xsyw, mod_ii, 
          mod_xsyw, mod_xsyw_lll, mod_xsyw_log, res_lll, stp_ii, stp_xsyw, stp_xsyw_lll, 
          stp_xsyw_log, test_ii, test_log, test_shapiro, test_xsyw, train_ii, train_log, 
          train_xsyw, vif_ii, xsyw, xxB27Hdesc, xxB27Hmod, xxB27Hraw)
```

```{r 'B27-Validation', include=FALSE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
