# Logistic Regression (B34, Feb-27) {#b34 .unlisted .unnumbered}

```{r 'B34', include=FALSE, cache=FALSE, eval=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```


## Data: Bank {#set-bank-b34 .tabset .tabset-fade}

### Glance {.unlisted .unnumbered}

\textcolor{pink}{Please import the "B34-bank-full.csv"}

- About: Bank [45211, 17]
  - 'y' is the Response Variable: has the client subscribed a term deposit
    - Goal is to estimate /predict if the client will subscribe a bank term deposit
  - default: has credit in default (categorical: "no", "yes")
  - balance: Its description is not known yet - "ForLater"
    - It can be negative also
  - housing: has housing loan (categorical: "no", "yes")
  - loan: has personal loan (categorical: "no", "yes")
  - related with the last contact of the current campaign:
    - day: It looks like the date of the month because [1, 31]
    - duration: last contact duration, in seconds (numeric). 
      - \textcolor{orange}{Caution:} This attribute highly affects the output target (e.g., if duration=0 then y="no"). Yet, the duration is not known before a call is performed. Also, after the end of the call y is obviously known. Thus, this input should only be included for benchmark purposes and should be discarded if the intention is to have a realistic predictive model.
  - other attributes:
    - campaign: number of contacts performed during this campaign and for this client (numeric, includes last contact)
    - pdays: number of days that passed by after the client was last contacted from a previous campaign (numeric; 999 means client was not previously contacted)
      - 36954 / 45211 values are "-1"
    - previous: number of contacts performed before this campaign and for this client (numeric)
      - For those 36954 records, this value is "0"
      - Maximum contacts performed for a client are 275 whereas 2nd highest is 58
    - poutcome: outcome of the previous marketing campaign (categorical)
      - It is "unknown" for those 36954 records and 5 more records
      - The "other" should be merged with "unknown"


```{r 'F67-Data-Bank', include=FALSE, eval=FALSE}
# #Path of Object, FileName and MD5Sum tools::md5sum(paste0(.z$XL, "B34-bank-full.csv"))
#xxB34Bank <- f_getObject("xxB34Bank", "B34-bank-full.csv",
                                #"e6b0ca77f3f200ec5428e04dd104da53")
loc_src <- paste0(.z$XL, "B34-bank-full", ".csv")
loc_rds <- paste0(.z$XL, "xxB34Bank", ".rds")
tbl <- read_delim(file = paste0(.z$XL, "B34-bank-full.csv"), delim = ";")
attr(tbl, "spec") <- NULL
attr(tbl, "problems") <- NULL
saveRDS(tbl, loc_rds)
```

```{r 'F67-Save-Bank', include=FALSE, eval=FALSE}
f_setRDS(xxB34Bank)
```

### EDA {.unlisted .unnumbered}

```{conjecture 'Tibble-Subset-NoMatrix'}
\textcolor{brown}{Error: The `i` argument of `[` ...}
```

- Cannot subset a Tibble using Matrix. Convert to a vector.

```{r 'F67-Get-Bank', include=FALSE}
xxB34Bank <- f_getRDS(xxB34Bank)
#""  "" ""
# #Factor | Relocate Response | Drop
xfw <- xxB34Bank %>% 
  #mutate(across(where(is.character) & !c(y), factor)) %>% 
  mutate(across(education, factor, 
                levels = c("primary", "secondary", "tertiary", "unknown"), 
                labels = c("I", "II", "III", "q"))) %>% 
  mutate(across(education, relevel, ref = "III")) %>% 
  rename(edu = education) %>% 
  mutate(across(marital, factor, 
                levels = c("married", "single", "divorced"), 
                labels = c("M", "S", "D"))) %>% 
  rename(wed = marital) %>% 
  mutate(across(contact, factor, 
                levels = c("cellular", "telephone", "unknown"), 
                labels = c("cell", "tele", "q"))) %>% 
  rename(call = contact) %>% 
  mutate(across(where(is.character), factor)) %>% 
  relocate(y) %>% 
  select(-c(job, month))
#
# #Create Dummies | Replaced All NA in original data so no replacement in dummy columns |
dum_xfw <- dummy_cols(xfw, 
  select_columns = c("wed", "edu", "default", "housing", "loan", "call", "poutcome"),
                      remove_first_dummy = TRUE, remove_selected_columns = TRUE) 
#
# #Partition Data
set.seed(3)
#idx_xsyw <- sample.int(n = nrow(dum_xfw), size = floor(0.8 * nrow(dum_xfw)), replace = FALSE)
idx_xsyw <- createDataPartition(xfw$y, p = 0.8, list = FALSE)
train_xfw <- dum_xfw[idx_xsyw[, 1], ] 
test_xfw  <- dum_xfw[-idx_xsyw[, 1], ]
#
xbank <- train_xfw
test_bank <- test_xfw
```


### Structure {.unlisted .unnumbered}

```{r 'F67-Structure-Bank'}
str(xbank)
```

### Summary {.unlisted .unnumbered}

```{r 'F67-Summary-Bank'}
summary(xbank)
```


### ETC {.unlisted .unnumbered}

```{r 'F67-ETC', include=TRUE, eval=FALSE}
bb <- xfw
# #Count NA in Columns
if(FALSE) colSums(is.na(bb)) %>% as_tibble(rownames = "Cols") %>% filter(value > 0)
# #Subset Rows
if(FALSE) bb %>% select(1) %>% slice(1:10)
# #Comma separated string having each item within quotes for easy pasting as character not objects
if(FALSE) cat('"', paste0(names(which(sapply(bb, is.factor))), collapse = '", "'), '"\n', sep = '')
if(FALSE) cat('"', paste0(levels(bb$Arrival), collapse = '", "'), '"\n', sep = '')
# #Filter
if(FALSE) bb %>% filter(is.na(bpl)) %>% select(id, bph, bpl)
# #Count Yes/No or True/False in ALL such Columns
if(FALSE) bb %>% select(iFemale, iMarried) %>% 
  pivot_longer(everything()) %>% count(name, value) %>% 
  pivot_wider(names_from = value, values_from = n) 
# #Count Unique of all Columns to decide which should be Factors
if(FALSE) bb %>% summarise(across(everything(), ~ length(unique(.)))) %>% pivot_longer(everything())
# #Summary of Columns of a class: is.factor is.numeric is.character
if(FALSE) bb %>% select(where(is.factor)) %>% summary()
# #Names and Indices of Columns of class: is.factor is.numeric is.character
if(FALSE) which(sapply(bb, is.factor))
# #Levels of Factor Columns
if(TRUE) lapply(bb["poutcome"], levels)
if(FALSE) lapply(bb[c(3)], levels)
# #Frequency of Each level of Factor
if(TRUE) bb %>% count(poutcome) %>% arrange(desc(n))
# #Coding for Dummy Variables
if(FALSE) contrasts(bb$Married) 
```



## Classification

```{r 'B34D01', comment="", echo=FALSE, results='asis'}
f_getDef("Classification") #dddd
```



## PERFORM LOGISTIC REGRESSION


```{r 'B34-T3', eval=FALSE}
mod_xfw_glm <- glm(y ~ ., family = binomial, data = train_xfw)
```

```{r 'B34-Print-Default-1', echo=FALSE, collapse=FALSE, class.output="models", eval=FALSE}
# #Coefficient Estimates, RSE, R2, F-statistic, Significance, DF
ttl_hh <- "Logistic Regression: Bank: Y ~ X"
col_hh <- c("ALL")
#
stargazer(mod_xfw_glm, 
    title = ttl_hh, column.labels = col_hh, model.numbers = FALSE, df = FALSE, report = "vc*", 
    type = "text", single.row = TRUE, intercept.bottom = FALSE, dep.var.caption = "", digits = 4)
```



```{r 'B34-T4', eval=FALSE}
#summary(mod_xfw_glm)
stargazer(mod_xfw_glm, type = "text", single.row = TRUE, intercept.bottom = FALSE)
```

## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'B34-Cleanup', include=FALSE, cache=FALSE, eval=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll)
```

```{r 'B34-Validation', include=FALSE, cache=FALSE, eval=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
