# MWE (Z01) {#z01}

```{r 'Z01', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## Overview

- Minimum Working Examples (MWE) for the most common /interesting /obscure 'how to do in R'.
  - It does not cover any function extensively or their finer points, cautions etc.
  - It assumes that reader knows R basics.

- [How to Load Packages](#loadpkg-z01 "z01")
- [How to Copy Clipboard as Tibble](#loaddata-z01 "z01")
- [How to Generate correlated vectors of normal distribution](#cor-vec-z01 "z01")
- [How to Count Unique Values in ALL Columns ](#n-uniq-z01  "z01")

## Load Multiple Packages {#loadpkg-z01 .tabset .tabset-fade}

### Packages {.unlisted .unnumbered}

```{r 'Z01-LoadPackages', eval=FALSE}
if(FALSE) {# #Load Packages that were already installed
  # #Install Packages if: Error in library(...) : there is no package called '...'
  # #WARNING: Installation may take some time.
  #install.packages("dplyr", dependencies = TRUE)
  library(dplyr)
  library(readr)
  library(tibble)
}
```

### ALL {.unlisted .unnumbered}

```{r 'Z01-LoadPackagesALL', eval=FALSE}
if(FALSE) {# #Load Packages that were already installed
  # #Install Packages if: Error in library(...) : there is no package called '...'
  # #WARNING: Installation may take some time.
  #install.packages("dplyr", dependencies = TRUE)
  pkg <- c("dplyr", "readr", "tibble")
  lapply(pkg, FUN = function(x) {suppressMessages(library(x, character.only = TRUE))})
  rm(pkg)
}
```

## Load Data {#loaddata-z01 .tabset .tabset-fade}

### Clipboard {.unlisted .unnumbered}

```{r 'Z01-LoadClipBoard', eval=FALSE}
if(FALSE) {
  library(readr)
}
# #Assuming that the Data has been copied to clipboard. Output is a Tibble.
ii <- read_delim(clipboard())
```


## Generate Correlated Vectors {#cor-vec-z01 .tabset .tabset-fade}

```{r 'Z01-CorrelatedVectors'}
# #Set Seed
set.seed(3)
#
# #Generate correlated vectors of normal distribution
ii <- rnorm(n = 50, mean = 0, sd = 1)
jj <- ii + rnorm(n = 50, mean = 1, sd = 0.1)
#
# #Check Correlation
cor(ii, jj)
```

## Count Unique Values in ALL Columns {#n-uniq-z01 .tabset .tabset-fade}

```{r 'Z01-CountUnique'}
# #Count Unique Values in ALL Columns
sapply(mtcars, function(x) length(unique(x)))
```


## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'Z01-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll)
```

```{r 'Z01-Validation', include=TRUE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
