# MWE (Z01) {#z01}

```{r 'Z01', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## Overview

- Minimum Working Examples (MWE) for the most common /interesting /obscure 'how to do in R'.
  - It does not cover any function extensively or their finer points, cautions etc.
  - It assumes that reader knows R basics.

- [How to Limit the Knit or Build](#limit-knit-z01 "z01")
- [How to Load Packages](#loadpkg-z01 "z01")
- [How to Copy Clipboard as Tibble](#load-data-z01 "z01")
- [How to provide Absolute Path or Relative Path](#load-data-z01 "z01")
- [How to Generate correlated vectors of normal distribution](#cor-vec-z01 "z01")
- [How to Count Unique Values in ALL Columns](#n-uniq-z01 "z01")
- [How to Rename ALL Headers to Proper Names](#rename-z01 "z01")
- [How to Merge Factor Levels](#merge-factor-level-z01 "z01")
- [How to Create a Contingency Table](#table-z01 "z01")
- [How to Add Rows and Column Totals to a Table or Confusion Matrix](#tab-sum-z01 "z01")
- [How to Check R Version](#version-z01 "z01")
- [How to Pipe the Subset](#pipe-sub-z01 "z01")
- [How to Test Background Colours](#bg-color-z01 "z01")
- [How to Modify Tibble Options](#mod-tbl-z01 "z01")
- [How to Change Colour Opacity](#col-alpha-z01 "z01")
- [How to Check if a Package is in the Installed Packages](#find-intall-package-z01 "z01")
- [DataFrame and Matrix Coercion](#df-mat-z01 "z01")


## Limit the Knit or Build {#limit-knit-z01 .tabset .tabset-fade}

```{r 'Z01-Limit-Knit-Build', eval=FALSE}
# #To only evaluate up to this chunk and complete the knit process.
knit_exit()
```

## Load Multiple Packages {#loadpkg-z01 .tabset .tabset-fade}

### Packages {.unlisted .unnumbered}

```{r 'Z01-LoadPackages', eval=FALSE}
if(FALSE) {# Install Missing Packages: Error in library(...) : there is no package called '...'
 install.packages("conflicted", dependencies = TRUE)
}

if(FALSE) {# #Load Packages that were already installed
  library("conflicted")
  #library("readr") #read_csv
  library("readxl") #read_excel
  library("dplyr")  #select
  library("tibble") #tibble
  library("tidyr")  #pivot_longer
  library("caret")  #createDataPartition
  library("ROSE")   #ovun.sample
  library("broom")  #augment
  library("stargazer")     #stargazer
  library("car")    #vif
  library("rpart")  #rpart
  library("randomForest")  #randomForest
  
  # Precedence
  # dplyr::filter | stats::filter
  conflict_prefer("filter", "dplyr")
}
```

### ALL {.unlisted .unnumbered}

```{r 'Z01-LoadPackagesALL', eval=FALSE}
if(FALSE) {# #Load Packages that were already installed
  # #Install Packages if: Error in library(...) : there is no package called '...'
  # #WARNING: Installation may take some time.
  #install.packages("dplyr", dependencies = TRUE)
  pkg <- c("dplyr", "readr", "tibble")
  lapply(pkg, FUN = function(x) {suppressMessages(library(x, character.only = TRUE))})
  rm(pkg)
}
```

## Load Data {#load-data-z01}

### Clipboard {.unlisted .unnumbered}

```{r 'Z01-LoadClipBoard', eval=FALSE}
if(FALSE) {
  library(readr)
}
# #Assuming that the Data has been copied to clipboard. Output is a Tibble.
ii <- read_delim(clipboard())
```

### PATH {.unlisted .unnumbered}

```{r 'Z01-PATH'}
# #[3] Conditions: A:Console, B: Manual R Markdown Chunk, C: Knit
loc_src <- "B09_xx_data.rds"
#
if(file.exists(paste0("./data/", loc_src))){
  # #Does not give Warning. However, it does not work during Manual Chunk evaluation i.e. AC not B
  file.exists(paste0("./data/", loc_src))
} else { #Works i.e. ABC but Warning with knitr::include_graphics()
  # #...It is highly recommended to use relative paths for images. You had absolute paths:
  file.exists(here("data", loc_src))
  file.exists(paste0(.z$XL, loc_src))
}
```

```{r 'Z01-Images', eval=FALSE}
# #Images 
loc_png <- paste0("./images/", "B38P01", "-Churn-Train-ROC", ".png")
if(FALSE){#Include images
  if(file.exists(loc_png)) include_graphics(loc_png) else include_graphics(here(loc_png))
}
if(FALSE){#Generate images
  if(!file.exists(here(loc_png))) {
    png(filename = here(loc_png))
    }
}
```

```{r 'Z01-Data, eval=FALSE'}
# #[3] Conditions: A:Console, B: Manual R Markdown Chunk, C: Knit
bb <- NA
loc_src <- here("data", "B09_xx_data.rds")  #ABC, No Warning after Knitr 1.39
if(FALSE){#AC, ERROR B
  loc_src <- paste0("./data/", "B09_xx_data.rds") 
  loc_src <- file.path(".", "data", "B09_xx_data.rds") #AC, ERROR B
}
file.exists(loc_src)
bb <- readRDS(loc_src)
bb
```


### R Basic PATHS {.unlisted .unnumbered}

- Use \textcolor{pink}{normalizePath()} to convert to proper path from DOS type
- Use \textcolor{pink}{R.home()} to get the Base PATH

```{r 'Z01-R-Path'}
normalizePath(R.home(component = "home"), winslash = "/", mustWork = TRUE)
normalizePath(R.home(component = "bin"), winslash = "/", mustWork = TRUE)
normalizePath(R.home(component = "library"), winslash = "/", mustWork = TRUE)
normalizePath(R.home(component = "doc"), winslash = "/", mustWork = TRUE)
normalizePath(R.home(component = "etc"), winslash = "/", mustWork = TRUE)
normalizePath(R.home(component = "include"), winslash = "/", mustWork = TRUE)
normalizePath(R.home(component = "modules"), winslash = "/", mustWork = TRUE)
normalizePath(R.home(component = "share"), winslash = "/", mustWork = TRUE)
normalizePath(R.home(component = "src"), winslash = "/", mustWork = TRUE)
normalizePath(R.home(component = "Tcl"), winslash = "/", mustWork = TRUE)
normalizePath(R.home(component = "tests"), winslash = "/", mustWork = TRUE)
```

### Startup {.unlisted .unnumbered}

- There are two sorts of files used in startup: 
  - \textcolor{pink}{Environment files:} contain lists of environment variables to be set
  - \textcolor{pink}{Profile files:} contain R code
- \textcolor{pink}{file.path()} is faster than using \textcolor{pink}{paste0()} for PATHS
  - AND it automatically handles the platform dependent separator

```{r 'Z01-Get-Env'}
# # R Startup Steps are given in the 'Initialization at Start of an R Session'
if(FALSE) help(startup)

# #A Named Vector of ALL Environment Variables
if(FALSE) Sys.getenv()

# #[1] Environment Variables: Given in ALL_CAPS
# #The name of the site file is the one pointed to by: R_ENVIRON
Sys.getenv("R_ENVIRON")

# #if R_ENVIRON is unset (as shown):
# #Then 'R_HOME/etc/Renviron.site' is used. However this file does not exist by default.
if(identical(Sys.getenv("R_ENVIRON"), "")){
  loc <- file.path(Sys.getenv("R_HOME"), "etc", "Renviron.site") 
  if(FALSE) normalizePath(loc, winslash = "/", mustWork = TRUE)
  file.exists(loc) #FILE DOES NOT EXIST
}

# #The name of the user file can be specified by: R_ENVIRON_USER
Sys.getenv("R_ENVIRON_USER")

# #if R_ENVIRON_USER is unset (as shown):
# #Then, the files searched for are '.Renviron' in the current and then in the user home directory
if(identical(Sys.getenv("R_ENVIRON_USER"), "")){
  loc <- file.path(Sys.getenv("R_USER"), ".Renviron")
  if(FALSE) normalizePath(loc, winslash = "/", mustWork = TRUE)
  file.exists(loc)  #FILE DOES NOT EXIST
}

# #[2] Then R searches for the site-wide startup profile file
# #The path of this file is taken from: R_ENVIRON
Sys.getenv("R_PROFILE")

# #if R_PROFILE is unset (as shown):
# #Then 'R_HOME/etc/Rprofile.site' is used. However this file does not exist by default.
if(identical(Sys.getenv("R_PROFILE"), "")){
  loc <- file.path(Sys.getenv("R_HOME"), "etc", "Rprofile.site")
  if(FALSE) normalizePath(loc, winslash = "/", mustWork = TRUE)
  file.exists(loc) #FILE EXISTS
}

# #Then R searches for a user profile
# #The path of this file is taken from: R_PROFILE_USER 
Sys.getenv("R_PROFILE_USER")

# #if R_PROFILE_USER is unset (as shown):
# #Then, the files searched for are '.Rprofile' in the current and then in the user home directory
if(identical(Sys.getenv("R_PROFILE_USER"), "")){
  loc <- file.path(Sys.getenv("R_USER"), ".Rprofile")
  if(FALSE) normalizePath(loc, winslash = "/", mustWork = TRUE)
  file.exists(loc)  #FILE EXISTS (MODIFIED)
}

# #EnvVar 
if(FALSE) ?HOME

# #Specific Environment Variable
Sys.getenv("R_HOME") #The top-level directory of the R installation
Sys.getenv("HOME")   #User HOME directory
Sys.getenv("R_USER") #User HOME directory. Set by R.

# Tilde "~" is USER HOME, Dot "." is Current Working Directory
read.delim(file.path("~", ".Rprofile"), nrows = 1, header = FALSE)
read.delim(file.path(".", ".Rprofile"), nrows = 1, header = FALSE)
```


## Generate Correlated Vectors {#cor-vec-z01 .tabset .tabset-fade}

```{r 'Z01-CorrelatedVectors'}
# #Set Seed
set.seed(3)
#
# #Generate correlated vectors of normal distribution
ii <- rnorm(n = 50, mean = 0, sd = 1)
jj <- ii + rnorm(n = 50, mean = 1, sd = 0.1)
#
# #Check Correlation
cor(ii, jj)
```

## Count Unique Values in ALL Columns {#n-uniq-z01 .tabset .tabset-fade}

```{r 'Z01-CountUnique'}
# #Count Unique Values in ALL Columns
sapply(mtcars, function(x) length(unique(x)))
```

## Rename ALL Headers to Proper Names {#rename-z01 .tabset .tabset-fade}

```{r 'Z01-RenameHeaders'}
if(FALSE) {
  library(dplyr)
}
# #Rename to Proper Names | To Lower, Replace by Underscore | 
aa <- tibble("A_Underscore" = 1, "B.Dot" = 2, "C Space" = 3, "D-Dash" = 4, `E'apostrophe` = 5) #'
bb <- aa %>% rename_with(make.names) %>% 
  rename_with(~ tolower(gsub(".", "_", .x, fixed = TRUE))) 
```

## Merge Factor Levels {#merge-factor-level-z01 .tabset .tabset-fade}

- Refer [abc ](#ref-merge-factor-c34 "c34")


```{r 'Z01-Merge-Factor-Levels'}
# #Alphabetical sorting by default i.e. No = 1, Yes = 2 but we want Yes = 1, No = 2
# #"H" should be converted to NA
bb <- tibble(x = factor(c("Y", "Yes", "Y", "N", "No", "H", NA)))
bb$x
#
ii <- bb %>% mutate(across(x, factor, levels = c("Y", "Yes", "N", "No"), 
                           labels = c("Yes", "Yes", "No", "No"))) 
ii$x
```

## Create a Contingency Table {#table-z01 .tabset .tabset-fade}

```{r 'Z01-Table'}
# #Contingency Table
aa <- c("TP_A" = 3, "FP_B" = 13, "TN_C" = 7, "TN_D" = 17) #c(3, 13, 7, 17)
bb <- as.table(matrix(aa, ncol = 2, byrow = TRUE, 
                dimnames = list("Prediction" = c("Yes", "No"), "Reference" = c("Yes", "No"))))
bb
```


## Add Rows and Column Totals to a Table or Confusion Matrix {#tab-sum-z01 .tabset .tabset-fade}

```{r 'Z01-Table-Total-Rows-Column'}
# #Add Row & Column using Magrittr Pipe
ii <- bb %>% as_tibble() %>% 
  pivot_wider(names_from = Reference, values_from = n) %>% 
  rename(Prediction_Reference = 1) %>% 
  add_row(summarise(., across(1, ~"Total")), summarise(., across(where(is.numeric), sum))) %>% 
  mutate(SUM = rowSums(across(where(is.numeric))))
#
# #Add Row & Column using Base Pipe
jj <- bb |> as_tibble() |> 
  pivot_wider(names_from = Reference, values_from = n) |> 
  rename(Prediction_Reference = 1) |> 
  {\(x) add_row(x, summarise(x, across(1, ~"Total")), 
                summarise(x, across(where(is.numeric), sum)))}() |> 
  mutate(SUM = rowSums(across(where(is.numeric))))
stopifnot(identical(ii, jj))
jj
```


## Check R Version {#version-z01 .tabset .tabset-fade}

```{r 'Z01-R-Version'}
# #Version 4 has introduced Pipe "|>" in R.
R.version.string

if(TRUE) stopifnot(4L <= as.integer(R.version$major))
```

## Pipe the Subset {#pipe-sub-z01 .tabset .tabset-fade}

```{r 'Z01-Pipe-Subset'}
set.seed(3)
ii <- setNames(sample.int(n = 100, size = 10), nm = letters[1:10])
ii
# #Subset Numbers greater than 50
jj <- ii %>% `[`(. > 50)
kk <- ii |> {\(x) x[x > 50]}()
stopifnot(identical(jj, kk))
kk
```

## Test Background Colours {#bg-color-z01 .tabset .tabset-fade}

```{r 'Z01-BG-Colour-Models', collapse=FALSE, class.output="models"}
print("Models: Color= #80A080, background-color: #303030")
```

```{r 'Z01-BG-Warn-Red', collapse=FALSE, class.source='redwarnimp', class.output="redwarn"}
tryCatch(warning("WARNING: background-color: #CF6679FF (Red)"), 
         warning = function(w) {
           print(trimws(gsub("[^:]*:(.*)", "\\1", paste0(w))))
           #suppressWarnings() #Double Arrow Assignment
           }
         )
```

```{r 'Z01-BG-Primary', collapse=FALSE, class.output='bg-primary'}
# class.source works if '!important' override is given in the css
# Source Chunk generally have highlights so change in its color should be avoided
# Output chunk have a single coloured text, so its background can be modified safely
# class.output can be {'bg-primary', 'bg-success', 'bg-info', 'bg-warning', 'bg-danger'}
head(letters) #Primary
```

```{r 'Z01-BG-Success', echo=FALSE, collapse=FALSE, class.output='bg-success'}
head(letters) #Success
```

```{r 'Z01-BG-Info', echo=FALSE, collapse=FALSE, class.output='bg-info'}
head(letters) #Info
```

```{r 'Z01-BG-Warning', echo=FALSE, collapse=FALSE, class.output='bg-warning'}
head(letters) #Warning
```

```{r 'Z01-BG-Danger', echo=FALSE, collapse=FALSE, class.output='bg-danger'}
head(letters) #Danger
```


## Modify Tibble Options {#mod-tbl-z01 .tabset .tabset-fade}

### Rows {.unlisted .unnumbered}

```{r 'Z01-TibbleOptions-Rows'}
# #Change Number of Rows Printed
aa <- as_tibble(mtcars)
```

### Rows 20 {.unlisted .unnumbered}

```{r 'Z01-TibbleOptions-Rows20'}
# #Get Options: Default i.e. NULL means 20 Rows are Printed for a Tibble
getOption("pillar.print_max")
getOption("pillar.print_min")
#
head(aa, 20) #Prints 20 Rows - OK
head(aa, 21) #Prints 10 Rows - Surprise
print(aa, n = 21) #Prints 21 Rows - OK
```

### Rows 25 {.unlisted .unnumbered}

```{r 'Z01-TibbleOptions-Rows25'}
# #Modify Options
options(pillar.print_max = 25, pillar.print_min = 25)
head(aa, 25) #Prints 25 Rows - OK
# #Reset Options
options(pillar.print_max = NULL, pillar.print_min = NULL)
```

## Change Colour Opacity {#col-alpha-z01 .tabset .tabset-fade}

```{r 'Z01-R-Colour-Opacity'}
# #Vector of Color Hex with 100% (Last two characters are FF)
ii <- viridis_pal()(3)
ii
# #Split String and Modify and Paste (50% i.e. 7F)
jj <- paste0(substr(ii, 1, 7), "7", substr(ii, 9, 9))
jj
if(FALSE) {
  show_col(ii)
  show_col(jj)
}
```

## Check if a Package is in the Installed Packages {#find-intall-package-z01}

```{r 'Z01-Find-Installed-Pakage'}
# #Matrix of Installed Packages | Vector | To Lower for easy matching | Find Element |
installed.packages() |> rownames() |> tolower() |> is.element(el = "igraph")
```

## DataFrame and Matrix Coercion {#df-mat-z01}

```{r 'Z01-DF-Mat'}
aa_df <- data.frame(x = 1:4, y = 5:8, z = 9:12)
bb_mat <- matrix(1:12, ncol = 3, dimnames = list(NULL, c("x", "y", "z")))
# #Coercion
aa_mat_r <- as.matrix(aa_df)
bb_df_r <- as.data.frame(bb_mat, row.names = FALSE)
# #Verify
stopifnot(all(identical(aa_df, bb_df_r), identical(bb_mat, aa_mat_r)))
# #Matrix
aa_mat_r
# #DataFrame
bb_df_r
```

## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'Z01-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll, loc_src, loc)
```

```{r 'Z01-Validation', include=FALSE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
