# Advanced R (Z05) {#z05}

```{r 'Z05', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

- "ForLater" - "Advanced R", by Hadley Wickham, 2nd Edition, Chapman & Hall. 
  - Add the citation
  
## Names and values

### Binding Basics

- ~~"create an object named 'x', containing the values 11, 22, and 33"~~ is a simplification
  - There are two steps
    - It is creating an object, a vector of values, c(11, 22, 33).
    - And it is binding that object to a name, x.
  - In other words, the object, or value, does not have a name; it is actually the name that has a value.
  - Assignment Operator "<-" creates a binding from the name on the left-hand side (x) to the object on the right-hand side (vector containing values).
  - Thus, you can think of a name as a reference to a value. 

```{r 'Z05-22-Binding'}
# #x is the reference to the vector on the right side
# #NOTE: We are giving the right side object (without its binding) a label or identifier: "0x74b"
# #...It is analogous to its location in memory where the object is stored. But because ...
# #...the actual memory address changes every time the code is run, we use these labels instead.
x <- c(11, 22, 33)
#
# #We get another binding /reference to the same vector (not another object) when we do this:
y <- x
```


```{r 'Z05-22-GetAddress'}
# #Get Memory Address: Both x and y points to the same identifier
obj_addr(x)
obj_addr(y)
```

#### Syntactic Names

> A syntactic name must consist of letters, digits, dot (.) and underscore "_" but cannot begin with _ or a digit. Further, reserved words are not allowed. These rules can be overriden by using backticks (``)

```{r 'Z05-22-Reserved-Words'}
# #List of Reserved Words
if(FALSE) ?Reserved
```

#### Exercise

- Explain the relationship between aa, bb, cc and dd in the following code:

```{r 'Z05-22-Q1'}
aa <- 1:10
bb <- aa
cc <- bb
dd <- 1:10
#
# #a, b, c points to same object (same memory address). 
# #d points to different object which incidentally has same values.
stopifnot(all(identical(aa, bb), identical(aa, cc)))
obj_addr(aa)
obj_addr(dd)
```

- The following code accesses the 'mean' function in multiple ways. Do they all point to the same underlying function object
  - mean and bass::mean are pointing to different memory address (if there are other packages present)
  - they would point to same memory in a new R session

```{r 'Z05-22-Q2'}
# xxxx
if(FALSE) stopifnot(all(identical(obj_addr(mean), obj_addr(get("mean"))), 
              identical(obj_addr(mean), obj_addr(evalq(mean))),
              identical(obj_addr(mean), obj_addr(match.fun("mean")))))
#
obj_addr(mean)
obj_addr(get("mean"))
obj_addr(evalq(mean))
obj_addr(match.fun("mean"))
if(FALSE){# #Output remains same (as Hadley intended), if there are fewer packages loaded.
  library(lobstr)
  mean
#function (x, ...) 
#UseMethod("mean")
#<bytecode: 0x000002353cd4fe98>
#<environment: namespace:base>
  base::mean
#function (x, ...) 
#UseMethod("mean")
#<bytecode: 0x000002353cd4fe98>
#<environment: namespace:base>
  stopifnot(identical(obj_addr(mean), obj_addr(base::mean)))
  obj_addr(mean)
}

if(FALSE) {# #Output is different, if there are some more packages loaded.
  mean
#standardGeneric for "mean" defined from package "base"
#function (x, ...) 
#standardGeneric("mean")
#<environment: 0x0000020bf0a3abe8>
#Methods may be defined for arguments: x
#Use  showMethods(mean)  for currently available ones.
  base::mean
#function (x, ...) 
#UseMethod("mean")
#<bytecode: 0x0000020bdbcf5508>
#<environment: namespace:base>
}
```

#### Aside: More details on Memory

- [(SO) This](https://stackoverflow.com/questions/61080323 "https://stackoverflow.com")
- Any R object is a C (pointer, \textcolor{pink}{SEXP}, to a) "multi-object" (struct). This includes information about the R object and, also, the actual data of the R object that we have access to. Information consists of data that R needs to operate, e.g. length, number of references, to know when to copy an object etc.
- \textcolor{pink}{lobstr::obj_addr()}, presumably, returns the memory address that a SEXP points to. That part of the memory contains both the information about and the data of the R object. From within the R environment we cannot or do not access the (pointer to the) memory of the actual data in each R object.
- The function `[ ]` copies the requested element of the data contained in the C object to a new C object and returns its SEXP pointer to R. Each time the function is called, a new C object is created and returned to R.
- We cannot access the memory address of each element of the actual data of our object through R. However, we can trace the respective addresses using the C api.

```{r 'Z05-MemoryAddress'}
x <- c(11L, 22L, 33L)
y <- x
#
stopifnot(identical(obj_addr(x), obj_addr(y))) #No issue till here
obj_addr(y)
#
# #However, the address of elements look different 
obj_addr(x[1])
obj_addr(y[1])
#
# #A function to get the addresses:
z05_addr <- inline::cfunction(sig = c(x = "integer"), body = '
             Rprintf("SEXP @ %p\\n", x);

             Rprintf("first element of SEXP actual data @ %p\\n", INTEGER(x));

             for(int i = 0; i < LENGTH(x); i++) 
                 Rprintf("<%d> @ %p\\n", INTEGER(x)[i], INTEGER(x) + i);

             return(R_NilValue);
     ')
#
# #Check Memory Addresses: Addresses of vector values referred by x and y are same
if(FALSE){
  z05_addr(x)
  z05_addr(y) 
}
#
# #The successive memory difference between data elements equals the size of int type:
if(FALSE){# #Change here the new memory locations of x1, x2, x3
  # #NOTE: I was unable to execute this code. - "ForLater"
  diff(c(strtoi("0x00000257a7690360", 16L), 
       strtoi("0x00000257a7690364", 16L), 
       strtoi("0x00000257a7690368", 16L)))
}
#
# #However, sub-setting i.e. using [] leads to ...
# #We are running a function which returns a value. That value is an integer. 
# #We are evaluating the function x[1] and then giving you the address of what that function return
# #...not the address of the first element of the array x.
# #Trying to access same item results in different address every time
z05_addr(x[1])
z05_addr(x[1])
z05_addr(`[`(x, 1)) #]
z05_addr(`[`(x, 1)) #]
#
rm(z05_addr)
```

### Copy-on-modify

- When the value associated with y changes, the original object is not modified. 
- Instead, R creates a new object, 0xcd2, a copy of 0x74b with the modification, then rebinds y to that object.
- ~~R objects are unchangeable, or immutable.~~ There are some important exception.

```{r 'Z05-23-Copy-on-modify'}
# #x "0x74b" 
x <- c(11, 22, 33)
y <- x
#
obj_addr(x) #"0x74b"
obj_addr(y) #"0x74b"
#
# #y "0xcd2" : New Object with one value changed and rebind y to it.
y[[3]] <- 44
#
obj_addr(x) #"0x74b"
obj_addr(y) #"0xcd2" - New
#
print(list(x = x, y = y))
```

#### tracemem()

- \textcolor{pink}{tracemem()}
  - We will get the current address of the object and from then onwards, whenever that object is copied, it will print a message with the information of which object was copied, its new address, and the sequence of calls that led to the copy.
  - \textcolor{orange}{Caution:} Using tracemem in RStudio will result in additional copies because of the reference from the environment pane.

```{r 'Z05-23-tracemem'}
x <- c(11, 22, 33)
z <- x #Before Starting Trace
# #Start Memory Tracing
tracemem(x)
y <- x #After Starting Trace
#

# #Modify
y[[3]] <- 44 #Trace shown 
x[[3]] <- 55 #Trace shown 
#
##xxxx : check after knit if a -> b step is present or not
z[[3]] <- 66 #No Trace shown in R or RStudio Console. However, additional copy step found in Chunk
#
# #Turn off the memory tracing
untracemem(x); untracemem(y)
#
print(list(x = x, y = y, z = z))
```

#### Function Calls

- Similarly
  - While f() is running, the a inside the function points to the same value as the x does outside the function.
  - the function f() has a formal argument, a, which becomes a binding in the execution environment when the function is run.
  - Once f() completes, x and z will point to the same object. 0x74b never gets copied because it never gets modified. 
  - If f() did modify x, R would create a new copy, and then z would bind that object.

```{r 'Z05-23-tracemem-function'}
x <- c(11, 22, 33)
tracemem(x)
#
# #Function: f() 
z05_f <- function(a) {
  a
}
#
# #No Copy
z <- z05_f(x)
#
untracemem(x)
#
identical(obj_addr(x), obj_addr(z))
obj_addr(x)
print(list(x = x, z = z))
rm(z05_f)
```

#### Lists

- List is more complex because instead of storing the values itself, it stores references to them.
- This is a \textcolor{pink}{shallow copy}: the list object and its bindings are copied, but the values pointed to by the bindings are not. 
  - The opposite of a shallow copy is a \textcolor{pink}{deep copy} where the contents of every reference are copied. 
  - Before version 3.1.0 R was always producing deep copies.
  - This was the reason for original memory heavy performance of R during copy.

- \textcolor{pink}{ref()}
  - It prints the memory address of each object, along with a local ID so that you can easily cross-reference shared components.

```{r 'Z05-23-Lists'}
# #List l1 points to addresses pointing each to the values
l1 <- list(11, 22, 33)
l2 <- l1
#
# #Now, List l2 points to addresses pointing to first and second addresses (as original) 
# #...but third is pointing to a different address pointing to the new value.
l2[[3]] <- 44
#
# #ref() to compare addresses of addresses contained.
ref(l1, l2)
```

#### Data frames

- Data frames are lists of vectors
  - If you modify a column, only that column needs to be modified; the others will still point to their original references.
  - However, if you modify a row, every column is modified, which means every column must be copied.

```{r 'Z05-23-DataFrame'}
# #DataFrame
d1 <- data.frame(x = c(1, 2, 3), y = c(4, 5, 6))
d3 <- d2 <- d1
#
# #Modify Column: Address of only modified column is changed.
d2[ , 2] <- d2[ , 2] * 2
ref(d1, d2)
#
# #Modify Row: All Column References modified
d3[1, ] <- d3[1, ] * 3
ref(d1, d3)
```

#### Character vectors

- R uses a \textcolor{pink}{global string pool} where each element of a character vector is a pointer to a unique string in the pool

```{r 'Z05-23-Character'}
# #Note: both 1st and 2nd addresses are pointing to same value in the Global String Pool.
x <- c("a", "a", "abc", "d")
#
ref(x, character = TRUE)
```


#### Exercise

- Why is tracemem(1:10) not useful
  - At runtime, an object with an address in memory is created, but it is not bound to a name. Therefore, the object cannot be called or manipulated from R. As no copies will be made, it is not useful to track the object for copying.

```{r 'Z05-23-Q1'}
tracemem(1:10)
#
tracemem(1:10)
```

```{r 'Z05-23-Q2'}
# #Explain why tracemem() shows two copies when you run this code.
y <- x <- c(11L, 22L, 33L)
tracemem(x)
#
x[[3]] <- 44
#
# #The above has two steps: integer to double conversion and value modification
# #The change below shows only 1 step: valye modification
tracemem(y)
#
##xxxx : check after knit if a -> b step is present or not
y[[3]] <- 44L #No Trace shown in R or RStudio Console. However, additional copy step found in Chunk
#
untracemem(x); untracemem(y)
```


```{r 'Z05-23-Q3'}
x <- 1:10
y <- list(x, x)
z <- list(y, x, 1:10)
#
# #Relationship
# # z : an address having list of 3 addresses:
# #   First: an address having a list of 2 addresses: Both points to same address having values 1:10
# #  Second: an address having values 1:10, can be copied, tracemem works
# #   Third: an address having values 1:10, it cannot be copied, no tracemem
ref(z, character = TRUE)
z
```

```{r 'Z05-23-Q4'}
# #What happens here:
x <- list(1:10)
tracemem(x)
#
# #The initial reference tree of x shows that the name x binds to a list object. 
# #This object contains a reference to the integer vector 1:10.
ref(x, character = TRUE)
#
x[[2]] <- x
#
# #When x is assigned to an element of itself the list is copied to a new address in memory.
# #The list object previously bound to x is now referenced in the newly created list object. 
# #It is no longer bound to a name. The integer vector is referenced twice.
ref(x, character = TRUE)
#
x
```

### Object size

- \textcolor{pink}{obj_size()}
  - how much memory an object uses

- Since the elements of lists are references to values, the size of a list is generally smaller than expected
- Similarly, because R uses a global string pool, character vectors take up less memory than expected
- Alternative Representation: 
  - Instead of storing every number in sequence, R stores only first & last number.
  - This means that every sequence, no matter how large, is the same size.

```{r 'Z05-24-Size'}
obj_size(letters)
#
x <- runif(1e2)
obj_size(x)
y <- list(x, x, x)
obj_size(y)
#
# #Because y stores references, it is essentially additional size of list of 3 elements (80B)
obj_size(y) - obj_size(x)
obj_size(list(NULL, NULL, NULL))
#
# #Whereas this stores addresses of 3 elements (80B) and 3 objects each having same size of 848B
obj_size(list(runif(1e2), runif(1e2), runif(1e2))) 
2624L - 848L * 3 - 80L
#
# #ALTREP: All have same size
obj_size(1:3)
obj_size(1:1e6)
obj_size(1:1e9)
```

### Modify-in-place

- Modifying an R object usually creates a copy. There are two \textcolor{pink}{exceptions:}
  - Objects with a single binding get a special performance optimization.
    - When it comes to bindings, R can currently only count 0, 1, or many. That means that if an object has two bindings, and one goes away, the reference count does not go back to 1: one less than many is still many. In turn, this means that R will make copies when it sometimes does not need to.
    - \textcolor{orange}{Caution:} Unable to replicate this. x without any copies is also showing different address after modification.
    - Whenever you call the vast majority of functions, it makes a reference to the object. The only exception are specially written "primitive" C functions. These can only be written by R-core and occur mostly in the base package.
  - Together, these two complications make it hard to predict whether or not a copy will occur. 
  - Instead, it is better to determine it empirically with tracemem().
  - Modifying a list uses internal C code, so the references are not incremented and only a single copy is made.
  - Environments, a special type of object, are always modified in place.

```{r 'Z05-25-Single-Binding'}
x <- c(11, 22, 33) #0 Copies
y <- c(11, 22, 33) #1 Copy
yy <- y 
z <- c(11, 22, 33) #2 Copies
zz <- z
zzz <- z
#
tracemem(x)
tracemem(y)
tracemem(z)
#
ref(x, y, z, character = TRUE)
#
x[[3]] <- 44 #This memory address modification is a problem!
y[[3]] <- 44
z[[3]] <- 44
#
# #x gets modified in place i.e. keeps its place in memory whereas y, z had to be copied first
# #It was supposed to be done as described above but it is NOT working as expected. - "ForLater"
ref(x, y, z, character = TRUE)
#
untracemem(x); untracemem(y); untracemem(z)
```

#### Environments

- \textcolor{pink}{reference semantics} : Environments are always modified in place. 
  - Environments are one of the R object types that are \textcolor{pink}{mutable}.
  - When you modify an environment all existing bindings to that environment continue to have the same reference.
  - One consequence of this is that environments can contain themselves


```{r 'Z05-25-Environments'}
# #Create Environments
e1 <- rlang::env(x = 11, y = 22, z = 33)
e2 <- e1
#
# #Modification of an object in e1 changes value in the object of e2 also
e1$z <- 44
e1$z
e2$z
#
ref(e1$z, e2$z, character = TRUE)
#
# #Environment can contain itself
e <- rlang::env()
e$self <- e
#
ref(e, character = TRUE)
```

### Unbinding and the garbage collector

- \textcolor{pink}{gc()}
  - The garbage collector (GC) runs automatically whenever R needs more memory to create a new object.
  - There is never any need to call gc() yourself. 
  
#### Aside: NA & NaN


```{r 'Z05-NA-NaN'}
# #NA: Not available i.e. Missing Value 
NA | TRUE  #TRUE  #At least one TRUE is present
NA | FALSE #NA    #Depending upon what the unknown is, the outcome will change
NA & TRUE  #NA    #Depending upon what the unknown is, the outcome will change
NA & FALSE #FALSE #At least one FALSE is present
#
# #NaN: Same as NA. Output is NA not NaN
c(NaN | TRUE, NaN | FALSE, NaN & TRUE, NaN & FALSE)
#
# #Identity: Higher Precedence: y ^ 0 are 1: Anything raised to 0 is always 1 (even NaN)
c(0^0, NA^0, NaN^0, Inf^0)
#
# #Identity: Higher Precedence: 1 ^ y are 1 : 1 to the power of anything is always 1 
c(1^0, 1^NA, 1^NaN, 1^Inf)
#
# #Mathematical operation with NA is NA and NaN is NaN
c(0^NA, NA^NA, NaN^NA, Inf^NA)     #NA
c(0^NaN, NA^NaN, NaN^NaN, Inf^NaN) #NaN
#
# #This above described higher precedence does not exist for other operations
NA * 0   #NA
NaN * 0  #NaN
#
# #Other inconsistencies
prod(NA, na.rm = TRUE)  #Instead of NA, surprisingly returns 1 
#
# #Special Mention: Non-commutative addition and multiplication with NA & NaN
# #Thumb rule: Output is first or left item except in case of power where it is the base
c("ApB" = NA + NaN, "BpA" = NaN + NA, "AmB" = NA * NaN, "BmA" = NaN * NA, 
  "AdB" = NA / NaN, "BdA" = NaN / NA, "AcB" = NA^NaN, "BcA" = NaN^NA)
#
# #Missing values (NA) and NaN values are regarded as non-comparable even to themselves, ...
# #...so comparisons involving them will always result in NA. 
identical(NA, NA) #TRUE These are identical 
NA == NA          #...but not comparable, use is.na()
#
identical(NaN, NaN) #TRUE These are identical 
NaN == NaN          #...but not comparable, use is.nan()
```

#### Others

- Note that as.vector() and is.vector() use different definitions of "vector!"


```{r 'B-'}
rm(x)
x
x <- 1
x

```


## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'Z05-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll)
```

```{r 'Z05-Validation', include=FALSE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
