# Advanced R (Z05) {#z05}

```{r 'Z05', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

- "ForLater" - "Advanced R", by Hadley Wickham, 2nd Edition, Chapman & Hall. 
  - Add the citation
  
## Names and values

### Binding Basics

- ~~"create an object named 'x', containing the values 11, 22, and 33"~~ is a simplification
  - There are two steps
    - It is creating an object, a vector of values, c(11, 22, 33).
    - And it is binding that object to a name, x.
  - In other words, the object, or value, does not have a name; it is actually the name that has a value.
  - Assignment Operator "<-" creates a binding from the name on the left-hand side (x) to the object on the right-hand side (vector containing values).
  - Thus, you can think of a name as a reference to a value. 

```{r 'Z05-22-Binding'}
# #x is the reference to the vector on the right side
# #NOTE: We are giving the right side object (without its binding) a label or identifier: "0x74b"
# #...It is analogous to its location in memory where the object is stored. But because ...
# #...the actual memory address changes every time the code is run, we use these labels instead.
x <- c(11, 22, 33)
#
# #We get another binding /reference to the same vector (not another object) when we do this:
y <- x
```


```{r 'Z05-22-GetAddress'}
# #Get Memory Address: Both x and y points to the same identifier
obj_addr(x)
obj_addr(y)
```

#### Syntactic Names

> A syntactic name must consist of letters, digits, dot (.) and underscore "_" but cannot begin with _ or a digit. Further, reserved words are not allowed. These rules can be overriden by using backticks (``)

```{r 'Z05-22-Reserved-Words'}
# #List of Reserved Words
if(FALSE) ?Reserved
```

#### Exercise

- Explain the relationship between aa, bb, cc and dd in the following code:

```{r 'Z05-22-Q1'}
aa <- 1:10
bb <- aa
cc <- bb
dd <- 1:10
#
# #a, b, c points to same object (same memory address). 
# #d points to different object which incidentally has same values.
stopifnot(all(identical(aa, bb), identical(aa, cc)))
obj_addr(aa)
obj_addr(dd)
```

- The following code accesses the 'mean' function in multiple ways. Do they all point to the same underlying function object
  - mean and bass::mean are pointing to different memory address (if there are other packages present)
  - they would point to same memory in a new R session

```{r 'Z05-22-Q2'}
if(FALSE) stopifnot(all(identical(obj_addr(mean), obj_addr(get("mean"))), 
              identical(obj_addr(mean), obj_addr(evalq(mean))),
              identical(obj_addr(mean), obj_addr(match.fun("mean")))))
#
obj_addr(mean)
obj_addr(get("mean"))
obj_addr(evalq(mean))
obj_addr(match.fun("mean"))
if(FALSE){# #Output remains same (as Hadley intended), if there are fewer packages loaded.
  library(lobstr)
  mean
#function (x, ...) 
#UseMethod("mean")
#<bytecode: 0x000002353cd4fe98>
#<environment: namespace:base>
  base::mean
#function (x, ...) 
#UseMethod("mean")
#<bytecode: 0x000002353cd4fe98>
#<environment: namespace:base>
  stopifnot(identical(obj_addr(mean), obj_addr(base::mean)))
  obj_addr(mean)
}

if(FALSE) {# #Output is different, if there are some more packages loaded.
  mean
#standardGeneric for "mean" defined from package "base"
#function (x, ...) 
#standardGeneric("mean")
#<environment: 0x0000020bf0a3abe8>
#Methods may be defined for arguments: x
#Use  showMethods(mean)  for currently available ones.
  base::mean
#function (x, ...) 
#UseMethod("mean")
#<bytecode: 0x0000020bdbcf5508>
#<environment: namespace:base>
}
```

#### Aside: More details on Memory

- [(SO) This](https://stackoverflow.com/questions/61080323 "https://stackoverflow.com")
- Any R object is a C (pointer, \textcolor{pink}{SEXP}, to a) "multi-object" (struct). This includes information about the R object and, also, the actual data of the R object that we have access to. Information consists of data that R needs to operate, e.g. length, number of references, to know when to copy an object etc.
- \textcolor{pink}{lobstr::obj_addr()}, presumably, returns the memory address that a SEXP points to. That part of the memory contains both the information about and the data of the R object. From within the R environment we cannot or do not access the (pointer to the) memory of the actual data in each R object.
- The function `[ ]` copies the requested element of the data contained in the C object to a new C object and returns its SEXP pointer to R. Each time the function is called, a new C object is created and returned to R.
- We cannot access the memory address of each element of the actual data of our object through R. However, we can trace the respective addresses using the C api.

```{r 'Z05-MemoryAddress'}
x <- c(11L, 22L, 33L)
y <- x
#
stopifnot(identical(obj_addr(x), obj_addr(y))) #No issue till here
obj_addr(y)
#
# #However, the address of elements look different 
obj_addr(x[1])
obj_addr(y[1])
#
# #A function to get the addresses:
z05_addr <- inline::cfunction(sig = c(x = "integer"), body = '
             Rprintf("SEXP @ %p\\n", x);

             Rprintf("first element of SEXP actual data @ %p\\n", INTEGER(x));

             for(int i = 0; i < LENGTH(x); i++) 
                 Rprintf("<%d> @ %p\\n", INTEGER(x)[i], INTEGER(x) + i);

             return(R_NilValue);
     ')
#
# #Check Memory Addresses: Addresses of vector values referred by x and y are same
if(FALSE){
  z05_addr(x)
  z05_addr(y) 
}
#
# #The successive memory difference between data elements equals the size of int type:
if(FALSE){# #Change here the new memory locations of x1, x2, x3
  # #NOTE: I was unable to execute this code. - "ForLater"
  diff(c(strtoi("0x00000257a7690360", 16L), 
       strtoi("0x00000257a7690364", 16L), 
       strtoi("0x00000257a7690368", 16L)))
}
#
# #However, sub-setting i.e. using [] leads to ...
# #We are running a function which returns a value. That value is an integer. 
# #We are evaluating the function x[1] and then giving you the address of what that function return
# #...not the address of the first element of the array x.
# #Trying to access same item results in different address every time
z05_addr(x[1])
z05_addr(x[1])
z05_addr(`[`(x, 1)) #]
z05_addr(`[`(x, 1)) #]
#
rm(z05_addr)
```

### Copy-on-modify

- When the value associated with y changes, the original object is not modified. 
- Instead, R creates a new object, 0xcd2, a copy of 0x74b with the modification, then rebinds y to that object.
- ~~R objects are unchangeable, or immutable.~~ There are some important exception.

```{r 'Z05-23-Copy-on-modify'}
# #x "0x74b" 
x <- c(11, 22, 33)
y <- x
#
obj_addr(x) #"0x74b"
obj_addr(y) #"0x74b"
#
# #y "0xcd2" : New Object with one value changed and rebind y to it.
y[[3]] <- 44
#
obj_addr(x) #"0x74b"
obj_addr(y) #"0xcd2" - New
#
print(list(x = x, y = y))
```

#### tracemem()

- \textcolor{pink}{tracemem()}
  - We will get the current address of the object and from then onwards, whenever that object is copied, it will print a message with the information of which object was copied, its new address, and the sequence of calls that led to the copy.
  - \textcolor{orange}{Caution:} Using tracemem in RStudio will result in additional copies because of the reference from the environment pane.

```{r 'Z05-23-tracemem'}
x <- c(11, 22, 33)
z <- x #Before Starting Trace
# #Start Memory Tracing
tracemem(x)
y <- x #After Starting Trace
#
# #Modify
if(FALSE) y[[3]] <- 44 #Trace shown 
if(FALSE) x[[3]] <- 55 #Trace shown 
#
# #No Trace shown in R or RStudio Console. 
# #However, additional copy step (a -> b) was found in Chunk (manual or Knit)
if(FALSE) z[[3]] <- 66
#
# #Turn off the memory tracing
untracemem(x); untracemem(y)
#
print(list(x = x, y = y, z = z))
```

#### Function Calls

- Similarly
  - While f() is running, the a inside the function points to the same value as the x does outside the function.
  - the function f() has a formal argument, a, which becomes a binding in the execution environment when the function is run.
  - Once f() completes, x and z will point to the same object. 0x74b never gets copied because it never gets modified. 
  - If f() did modify x, R would create a new copy, and then z would bind that object.

```{r 'Z05-23-tracemem-function'}
x <- c(11, 22, 33)
tracemem(x)
#
# #Function: f() 
z05_f <- function(a) {
  a
}
#
# #No Copy
z <- z05_f(x)
#
untracemem(x)
#
identical(obj_addr(x), obj_addr(z))
obj_addr(x)
print(list(x = x, z = z))
rm(z05_f)
```

#### Lists

- List is more complex because instead of storing the values itself, it stores references to them.
- This is a \textcolor{pink}{shallow copy}: the list object and its bindings are copied, but the values pointed to by the bindings are not. 
  - The opposite of a shallow copy is a \textcolor{pink}{deep copy} where the contents of every reference are copied. 
  - Before version 3.1.0 R was always producing deep copies.
  - This was the reason for original memory heavy performance of R during copy.

- \textcolor{pink}{ref()}
  - It prints the memory address of each object, along with a local ID so that you can easily cross-reference shared components.

```{r 'Z05-23-Lists'}
# #List l1 points to addresses pointing each to the values
l1 <- list(11, 22, 33)
l2 <- l1
#
# #Now, List l2 points to addresses pointing to first and second addresses (as original) 
# #...but third is pointing to a different address pointing to the new value.
l2[[3]] <- 44
#
# #ref() to compare addresses of addresses contained.
ref(l1, l2)
```

#### Data frames

- Data frames are lists of vectors
  - If you modify a column, only that column needs to be modified; the others will still point to their original references.
  - However, if you modify a row, every column is modified, which means every column must be copied.

```{r 'Z05-23-DataFrame'}
# #DataFrame
d1 <- data.frame(x = c(1, 2, 3), y = c(4, 5, 6))
d3 <- d2 <- d1
#
# #Modify Column: Address of only modified column is changed.
d2[ , 2] <- d2[ , 2] * 2
ref(d1, d2)
#
# #Modify Row: All Column References modified
d3[1, ] <- d3[1, ] * 3
ref(d1, d3)
```

#### Character vectors

- R uses a \textcolor{pink}{global string pool} where each element of a character vector is a pointer to a unique string in the pool

```{r 'Z05-23-Character'}
# #Note: both 1st and 2nd addresses are pointing to same value in the Global String Pool.
x <- c("a", "a", "abc", "d")
#
ref(x, character = TRUE)
```


#### Exercise

- Why is tracemem(1:10) not useful
  - At runtime, an object with an address in memory is created, but it is not bound to a name. Therefore, the object cannot be called or manipulated from R. As no copies will be made, it is not useful to track the object for copying.

```{r 'Z05-23-Q1'}
tracemem(1:10)
#
tracemem(1:10)
```

```{r 'Z05-23-Q2'}
# #Explain why tracemem() shows two copies when you run this code.
y <- x <- c(11L, 22L, 33L)
tracemem(x)
#
if(FALSE) x[[3]] <- 44
#
# #The above has two steps: integer to double conversion and value modification
# #The change below shows only 1 step: valye modification
tracemem(y)
#
# #No Trace shown in R or RStudio Console. 
# #However, additional copy step (a -> b) was found in Chunk (manual or Knit)
if(FALSE) y[[3]] <- 44L
#
untracemem(x); untracemem(y)
```


```{r 'Z05-23-Q3'}
x <- 1:10
y <- list(x, x)
z <- list(y, x, 1:10)
#
# #Relationship
# # z : an address having list of 3 addresses:
# #   First: an address having a list of 2 addresses: Both points to same address having values 1:10
# #  Second: an address having values 1:10, can be copied, tracemem works
# #   Third: an address having values 1:10, it cannot be copied, no tracemem
ref(z, character = TRUE)
z
```

```{r 'Z05-23-Q4'}
# #What happens here:
x <- list(1:10)
if(FALSE) tracemem(x)
#
# #The initial reference tree of x shows that the name x binds to a list object. 
# #This object contains a reference to the integer vector 1:10.
if(FALSE) ref(x, character = TRUE)
#
x[[2]] <- x
#
# #When x is assigned to an element of itself the list is copied to a new address in memory.
# #The list object previously bound to x is now referenced in the newly created list object. 
# #It is no longer bound to a name. The integer vector is referenced twice.
if(FALSE) ref(x, character = TRUE)
#
x
```

### Object size

- \textcolor{pink}{obj_size()}
  - how much memory an object uses

- Since the elements of lists are references to values, the size of a list is generally smaller than expected
- Similarly, because R uses a global string pool, character vectors take up less memory than expected
- Alternative Representation: 
  - Instead of storing every number in sequence, R stores only first & last number.
  - This means that every sequence, no matter how large, is the same size.

```{r 'Z05-24-Size'}
obj_size(letters)
#
x <- runif(1e2)
obj_size(x)
y <- list(x, x, x)
obj_size(y)
#
# #Because y stores references, it is essentially additional size of list of 3 elements (80B)
obj_size(y) - obj_size(x)
obj_size(list(NULL, NULL, NULL))
#
# #Whereas this stores addresses of 3 elements (80B) and 3 objects each having same size of 848B
obj_size(list(runif(1e2), runif(1e2), runif(1e2))) 
2624L - 848L * 3 - 80L
#
# #ALTREP: All have same size
obj_size(1:3)
obj_size(1:1e6)
obj_size(1:1e9)
```

### Modify-in-place

- Modifying an R object usually creates a copy. There are two \textcolor{pink}{exceptions:}
  - Objects with a single binding get a special performance optimization.
    - When it comes to bindings, R can currently only count 0, 1, or many. That means that if an object has two bindings, and one goes away, the reference count does not go back to 1: one less than many is still many. In turn, this means that R will make copies when it sometimes does not need to.
    - \textcolor{orange}{Caution:} Unable to replicate this. x without any copies is also showing different address after modification.
    - Whenever you call the vast majority of functions, it makes a reference to the object. The only exception are specially written "primitive" C functions. These can only be written by R-core and occur mostly in the base package.
  - Together, these two complications make it hard to predict whether or not a copy will occur. 
  - Instead, it is better to determine it empirically with tracemem().
  - Modifying a list uses internal C code, so the references are not incremented and only a single copy is made.
  - Environments, a special type of object, are always modified in place.

```{r 'Z05-25-Single-Binding'}
x <- c(11, 22, 33) #0 Copies
y <- c(11, 22, 33) #1 Copy
yy <- y 
z <- c(11, 22, 33) #2 Copies
zz <- z
zzz <- z
#
tracemem(x)
tracemem(y)
tracemem(z)
#
ref(x, y, z, character = TRUE)
#
if(FALSE) x[[3]] <- 44 #This memory address modification is a problem!
if(FALSE) y[[3]] <- 44
if(FALSE) z[[3]] <- 44
#
# #x gets modified in place i.e. keeps its place in memory whereas y, z had to be copied first
# #It was supposed to be done as described above but it is NOT working as expected. - "ForLater"
ref(x, y, z, character = TRUE)
#
untracemem(x); untracemem(y); untracemem(z)
```

#### Environments

- \textcolor{pink}{reference semantics} : Environments are always modified in place. 
  - Environments are one of the R object types that are \textcolor{pink}{mutable}.
  - When you modify an environment all existing bindings to that environment continue to have the same reference.
  - One consequence of this is that environments can contain themselves


```{r 'Z05-25-Environments'}
# #Create Environments
e1 <- rlang::env(x = 11, y = 22, z = 33)
e2 <- e1
#
# #Modification of an object in e1 changes value in the object of e2 also
e1$z <- 44
e1$z
e2$z
#
ref(e1$z, e2$z, character = TRUE)
#
# #Environment can contain itself
e <- rlang::env()
e$self <- e
#
ref(e, character = TRUE)
```

### Unbinding and the garbage collector

- \textcolor{pink}{gc()}
  - The garbage collector (GC) runs automatically whenever R needs more memory to create a new object.
  - There is never any need to call gc() yourself. 
  
#### Aside: NA & NaN


```{r 'Z05-NA-NaN'}
# #NA: Not available i.e. Missing Value 
NA | TRUE  #TRUE  #At least one TRUE is present
NA | FALSE #NA    #Depending upon what the unknown is, the outcome will change
NA & TRUE  #NA    #Depending upon what the unknown is, the outcome will change
NA & FALSE #FALSE #At least one FALSE is present
#
# #NaN: Same as NA. Output is NA not NaN
c(NaN | TRUE, NaN | FALSE, NaN & TRUE, NaN & FALSE)
#
# #Identity: Higher Precedence: y ^ 0 are 1: Anything raised to 0 is always 1 (even NaN)
c(0^0, NA^0, NaN^0, Inf^0)
#
# #Identity: Higher Precedence: 1 ^ y are 1 : 1 to the power of anything is always 1 
c(1^0, 1^NA, 1^NaN, 1^Inf)
#
# #Mathematical operation with NA is NA and NaN is NaN
c(0^NA, NA^NA, NaN^NA, Inf^NA)     #NA
c(0^NaN, NA^NaN, NaN^NaN, Inf^NaN) #NaN
#
# #This above described higher precedence does not exist for other operations
NA * 0   #NA
NaN * 0  #NaN
#
# #Other inconsistencies
prod(NA, na.rm = TRUE)  #Instead of NA, surprisingly returns 1 
#
# #Special Mention: Non-commutative addition and multiplication with NA & NaN
# #Thumb rule: Output is first or left item except in case of power where it is the base
c("ApB" = NA + NaN, "BpA" = NaN + NA, "AmB" = NA * NaN, "BmA" = NaN * NA, 
  "AdB" = NA / NaN, "BdA" = NaN / NA, "AcB" = NA^NaN, "BcA" = NaN^NA)
#
# #Missing values (NA) and NaN values are regarded as non-comparable even to themselves, ...
# #...so comparisons involving them will always result in NA. 
identical(NA, NA) #TRUE These are identical 
NA == NA          #...but not comparable, use is.na()
#
identical(NaN, NaN) #TRUE These are identical 
NaN == NaN          #...but not comparable, use is.nan()
```

#### Others

- Note that as.vector() and is.vector() use different definitions of "vector!"

## Functional Programming


## Object-oriented Programming

- \textcolor{pink}{Polymorphism} means that a developer can consider a function interface separately from its implementation, making it possible to use the same function form for different types of input. 
- \textcolor{pink}{Encapsulation} means that the user does not need to worry about details of an object because they are encapsulated behind a standard interface.

- Ex: summary()
  - We can imagine summary() containing a series of if-else statements, but that would mean only the original author could add new implementations. 
  - An OOP system makes it possible for any developer to extend the interface with implementations for new types of input.

- OO systems call the type of an object its \textcolor{pink}{class}, and an implementation for a specific class is called a \textcolor{pink}{method}. 
  - A class defines what an object is and methods describe what that object can do. 
  - The class defines the \textcolor{pink}{fields}, the data possessed by every instance of that class. 
  - Classes are organised in a hierarchy so that if a method does not exist for one class, method of its parent is used, and the child is said to inherit behavior. 
    - For example, in R, an ordered factor inherits from a regular factor, and a generalized linear model inherits from a linear model. 
    - The process of finding the correct method given a class is called \textcolor{pink}{method dispatch}.
    
- In \textcolor{pink}{encapsulated} OOP, methods belong to objects or classes, and method calls typically look like object.method(arg1, arg2). This is called encapsulated because the object encapsulates both data (with fields) and behavior (with methods), and is the paradigm found in most popular languages.
- In \textcolor{pink}{functional} OOP, methods belong to generic functions, and method calls look like ordinary function calls: generic(object, arg2, arg3). This is called functional because from the outside it looks like a regular function call, and internally the components are also functions.


- Base R provides 3 OOP systems: S3, S4, and reference classes (RC):
  - \textcolor{pink}{S3} is an informal implementation of functional OOP and relies on common conventions rather than ironclad guarantees. This makes it easy to get started with.
  - \textcolor{pink}{S4} is a formal and rigorous rewrite of S3. It requires more upfront work than S3, but in return provides more guarantees and greater encapsulation. S4 is implemented in the base methods package, which is always installed with R.
  - \textcolor{pink}{RC} implements encapsulated OO. 
    - RC objects are a special type of S4 objects that are also mutable, i.e., instead of using usual copy-on-modify semantics, they can be modified in place. 
    - This makes them harder to reason about, but allows them to solve problems that are difficult to solve in the functional OOP style of S3 and S4.
    - \textcolor{pink}{R6} implements encapsulated OOP like RC, but resolves some important issues.

- The \textcolor{pink}{sloop} package provides helpers to explore this. 

### Base Types

- While everything is an object, not everything is object-oriented.
  - Base Object
  - OO Object - These have a 'class' attribute
  - Thus, \textcolor{pink}{class()} returns misleading results for Base, whereas \textcolor{pink}{s3_class()} returns consistent results
  - While only OO objects have a class attribute, every object has a base type given by \textcolor{pink}{typeof()}

```{r 'Z05-ObjectType'}
aa <- 1:10 
bb <- mtcars
#
# Base Object - No class attribute
is.object(aa)
otype(aa)
attr(aa, "class")
class(aa)       #Misleading results on Base Objects
s3_class(aa)
typeof(aa)

# OO Object
is.object(bb)
otype(bb)
attr(bb, "class")
class(bb)
s3_class(bb)
typeof(bb)
```

## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'Z05-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll, cc, d1, d2, d3, dd, e, e1, e2, l1, l2, x, y, yy, z, zz, zzz)
```

```{r 'Z05-Validation', include=FALSE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
