# RStudio Underscore (Z02) {#z02}

## The .Rprofile

- To open Help in Browser and not in RStudio Help Tab
  - `options(help_type = "html")` # open help in browser (using this)
  - `options(help_type = "text")` # open help internally 

## Modifications

- To Stop R Studio from opening help topics in their Tab and breaking all alignment
  - Open File as Administrator "C:\Program Files\RStudio\R\Options.R"
  - Comment Out the code for 'custom browseURL implementation'

```{r 'Z02-HelpTab', eval=FALSE}
# #2022 To Open Help Files in browser instead of RStudio Tab
# custom browseURL implementation.
#.rs.setOption("browser", function(url)
#{
#   .Call("rs_browseURL", url, PACKAGE = "(embedding)")
#})
```

## Plot Tab

- To Stop RStudio from opening Plots in the small Tab 
  - Open File as Administrator "C:\Program Files\RStudio\R\Tools.R"
  - Modify the code as shown
  
```{r 'Z02-Plots', eval=FALSE}
# #2022 To Open Plots in separate window instead of RStudio Tab
# set our graphics device as the default and cause it to be created/set
.rs.addFunction( "initGraphicsDevice", function()
{
   #options(device="RStudioGD")
   #grDevices::deviceIsInteractive("RStudioGD")
   grDevices::deviceIsInteractive()
})
```


****
