# RStudio Underscore (Z02) {#z02}

## The .Rprofile

- To open Help in Browser and not in RStudio Help Tab
  - `options(help_type = "html")` # open help in browser (using this)
  - `options(help_type = "text")` # open help internally 

## Modifications

- To Stop R Studio from opening help topics in their Tab and breaking all alignment
  - Open File as Administrator "C:\Program Files\RStudio\R\Options.R"
  - Comment Out the code for 'custom browseURL implementation'

```{r 'Z02-HelpTab', eval=FALSE}
# #2022 To Open Help Files in browser instead of RStudio Tab
# custom browseURL implementation.
#.rs.setOption("browser", function(url)
#{
#   .Call("rs_browseURL", url, PACKAGE = "(embedding)")
#})
```

## Plot Tab

- To Stop RStudio from opening Plots in the small Tab 
  - Open File as Administrator "C:\Program Files\RStudio\R\Tools.R"
  - Modify the code as shown
  
```{r 'Z02-Plots', eval=FALSE}
# #2022 To Open Plots in separate window instead of RStudio Tab
# set our graphics device as the default and cause it to be created/set
.rs.addFunction( "initGraphicsDevice", function()
{
   #options(device="RStudioGD")
   #grDevices::deviceIsInteractive("RStudioGD")
   grDevices::deviceIsInteractive()
})
```

## RStudio Tools | Global Options

- R Markdown
  - General | Basic
    - R Version: Default
    - Uncheck: 
      - Restore most recently opened project at startup
        - Enabling this leads to many packages in "loaded via a namespace (and not attached):"
        - After disable: only 2 packages show up: compiler (same as R), tools (extra)
        - Meaning of "loaded via a namespace (and not attached):"
          - It means the package (In this case R) can access the package functions/objects, but the user can not without explicitly loading the tools package where as stats, graphics, etc. are loaded and ready to go for the user.
      - Restore .RData into workspace at startup
      - Automatically notify me of updates to RStudio
      - Send automated crash reports to RStudio
    - Check: 
      - Restore previously open source documents at startup
      - Always save history (even when not saving .RData)
      - Remove duplicate entries in history
      - Wrap around when navigating to previous/next tab
    - Default working directory (when not in a project): ~
    - Save workspace to .RData on exit: Never
  - Code
    - Editing
      - Tab width: 1
      - Uncheck: 
        - Use native pipe operator, |> (requires R 4.1+)
        - Soft-wrap R source files
        - Continue comment when inserting new line
        - Focus console after executing from source
        - Enable code snippets
      - Check: 
        - Insert spaces for Tab
        - Auto-detect code indentation
        - Insert matching parenthesis /quotes
        - Auto-indent code after paste
        - Vertically align arguments in auto-indent
        - Enable hyperlink highlighting in editor
      - Surround selection on text insertion: Quotes & Brackets
      - Keybindings: Default
      - Ctrl+Enter executes: Multi-line R statement
    - Display
      - Check: 
        - Highlight selected word
        - Highlight selected line
        - Show line numbers
        - Show margin
        - Show indent guides
        - Blinking cursor
        - Highlight R function calls
        - Rainbow parentheses
      - Uncheck: 
        - Show whitespace characters
        - Allow drag and drop of text
        - Allow scroll past end of document
      - Margin column: 100
      - Fold Style: Start and End
    - Saving
      - Uncheck: 2
        - Strip trailing horizontal whitespace when saving
        - Automatically save when editor loses focus
      - Idle period: 10000ms
      - Default text encoding: Ask
      - Line ending conversion: Platform Native
    - Completion
      - Show code completions: Automatically
      - Show completions after characters entered: 1
      - Show completions after keyboard idls (ms): 1000
      - Uncheck: 3
        - Show help tooltip after function completions 
        - Show help tooltip on cursor idle
        - Use tab for multiple auto completions
    - Diagnostics
      - Keyboard idle time: 2000 ms
      - Uncheck: 3
        - Check arguments to R function calls
        - Warn if variable used has no definition in scope
        - Provide R style diagnostics ...
  - Console
    - Check:
      - Show syntax highlighting in console input
      - Different color for error or message output ...
      - Warn when automatic session suspension is paused
        - Number of seconds to delay warning: 5
    - Uncheck:
      - Limit visible console output ...
      - Automatically expand tracebacks in error inspector
      - Double-click to select words
    - Limit output line length to: 100
    - ANSI Escape Codes: Show ANSI colors
  - Appearance 
    - RStudio theme: Modern
    - Zoom: 90
    - Editor Font: Consolas
    - Editor font size: 10
    - Editor theme: Tomorrow Night Bright
  - Pane Layout
    - Top Left: Source
    - Bottom Left: Console
    - Top Right: Environment, Build
    - Bottom Right: Viewer, Presentation
  - Packages | Management
    - Primary CRAN repository: Global (CDN) - RStudio
      - CRAN repositories modified outside package preferences
    - Check: All 3 options
  - Basic
    - Check:
      - Show document outline ...
      - Soft-wrap R Markdown files
      - Show output inline for all R Markdown documents
      - Hide console automatically when executing notebook chunks
        - no idea what this does
    - Uncheck:
      - Execute setup chunk automatically in notebooks
        - Manually executing that because I was worried that it was slowing down the runtime of manual chunk execution. Although, No improvement observed with this change.
    - Show 
    - Show in document outline: Sections only
      - Chunks are available in Bottom Bar and adding them here, the outline becomes huge and unmanageable
    - Show output preview in: Window
    - Show equation and image previews: inline
    - Evaluate chunks in directory: Document
  - Advanced
    - Check:
      - Enable chunk background highlight: Helps in separating chunks and normal text
      - Show inline toolbar for R code chunks: 
        - Keep it for those rare occurrences when output of many chunks needs to be kept
    - Uncheck: Display render command in R Markdown tab
      - This might be the one which stops automatic open of final document after generation
  - Visual
    - Check:
      - Show document outline ...
      - Show margin column indicator ...
    - Editor content width (px): 700
    - Default spacing between list items: spaced
    - Automatic text wrapping (line breaks): none
    - Write references at the end of current: block
- Python
  - Uncheck "Automatically activate project-local Python environments"
    - Because, I am using external Python installation
    - Enabling this leads to automatic attachment of internal Python module and that is old version. Further, to attach my external Python module, I have to restart R
- Sweave
  - Typeset LaTeX into PDF using: pdfLaTeX
    - "ForLater" Unsure about changing it to XeLaTeX
  - 2 Checks
    - Clean auxiliary output after compile
    - Always enable Rnw concordance ...
- Spelling
  - Main Dictionary Language: English (United States)
  - Check all 3 options
  - "ForLater" Unable to made custom dictionaries work, so using User Dictionary
- Git /SVN
  - Check "Enable version control interface for RStudio projects"
  - Git Executable: C:\\Program Files\\Git\\bin\\git.exe
  - SSH RSA Key: ...\\.ssh\\id_rsa
- Terminal | General
  - Initial Directory: Project Directory
  - New terminals open with: Windows PowerShell
  - Check
    - Enable Python integration
    - All Others
  - Uncheck
    - Clickable web links
- Accessibility 
  - Check
    - Reduce user interface animations
    - Always show focus outlines (requires restart)


****
