# R (Z04) {#z04}

```{r 'Z04', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## library() vs. require() 

- Packages are loaded using \textcolor{pink}{library()}
  - Generally, `library("tm")` is preferable over `require("tm")`
  - If a package is NOT installed properly
    - It will show error when loaded by the library() function
    - It will NOT show error when loaded by the require() function. It will silently fail. Thus, require() must be used along with if() to check for FALSE. 
  - Do not use require() function, library() is better.

## Search Installed Packages 

```{r 'Z04-SearchInstalledPackages'}
# #Check if following Packages are already installed
pkg <- c("data.table", "ODBC", "abc", "xyz")
#
# #List those packages which are NOT installed
miss_pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
#
if(!identical(0L, length(miss_pkg))){
  cat("Install Following Packages: ", paste0(miss_pkg, collapse=", "), "\n")
}
```

## Remove Objects


```{r 'Z04-RemoveObjects', eval=FALSE}
if(FALSE){
  # #Remove All Including Hidden
  rm(list = ls(all.names=TRUE))
  # #Remove All but NOT the Hidden
  rm(list = ls())
  # #Remove All Except those listed
  rm(list = setdiff(ls(), c("a", "c")))
  # #Removing Matching "gg01_" 
  rm(list = ls(pattern = "gg\\d\\d_"))
  # #Remove objects from another Environment
  rm(ZZ, envir = .z)
}
```

## ERROR

### Generate {.unlisted .unnumbered}

```{r 'Z04-Error-Generate'}
# #try() creates Error Object which stores the 'error' as 'message'
ii <- try("x" + "y", silent = TRUE)
jj <- try(stopifnot(1 == 2), silent = TRUE)
#
# #Print ERROR as Message
# #message=FALSE : Output in 'render' tab when knitting and below the chunk, not at the build END.
# #message=TRUE  : Output in Chunk Result.
message(ii)
message(jj)
```

### tryCatch() {.unlisted .unnumbered}


```{r 'Z04-TryCatchError', eval=FALSE}
# #Error Handling using tryCatch()
tryCatch(expr = {
# #Error: A continuous variable can not be mapped to shape 
# #print() is needed inside script for ERROR capturing
    if(exists("rs_err")) print(rs_err) else print("GG Object Not Found")
  }, warning = function(w) {#Handle Warning
	print(paste0(w))
  }, error = function(e) {#Handle ERROR
# #without paste0(), print shows full Backtrace
    print(paste0(e))
  }, finally = {#Do this at the end before quitting the tryCatch structure
  })
# #Cleanup
rm(rs_err)
```


## Rprofile

### One Tilde


```{r 'Z04-Rprofile-One', eval=FALSE}
# #[1] Tilde "~/.Rprofile" # Sys.getenv("HOME") or Sys.getenv("R_USER") #Auto-Execution on Startup#

# #Modify Library Path
# #Note that only directories that exist at the time will be included.
# #It needs modified libPaths for accessing ALL Packages during New Startup. However...
# #Any change in libPaths here [1] gets overwritten later if another .Rprofile [2] is executed
# #So, these changes are repeated also in [2]
if (!identical(.libPaths()[1], "C:/Softwares/R/LibFiles")) {
  .libPaths( c( "C:/Softwares/R/LibFiles", .libPaths()) )
}

# #To Point to the Latest Project as working directory
loc <- trimws(readLines("D:/Analytics/WorkingDirectory.txt", n = 1))
if(file.exists(loc)) {
    setwd(loc) 
    rm(loc)
    # #Execute the .Rprofile [2] (No Auto-execution)
    if (file.exists(".Rprofile")) {
        source(".Rprofile")
    } else {
        message("'.Rprofile' does not exist in Directory: ", getwd())
    }
} else {
    message("Check 'WorkingDirectory.txt', Target Directory does not exist: ", loc)
    rm(loc)
}
```

### Two Dot


```{r 'Z04-Rprofile-Two', eval=FALSE}
# #[2] Dot "./.Rprofile" #NO Auto-Execution, sourced and run by [1] on Restart by "^+F10" "AHK F5"#

# [1] Tilde: Sys.getenv("HOME")
# [2] Dot  : Working Directory - Project specific .Rprofile under GIT management
# This file is copied by SetRProject.ps1 from D:\Analytics\PowerShell\RFiles

# #Set a CRAN mirror. Change the Target URL to the one having 'https' to avoid warnings.
local({r <- getOption("repos")
       r["CRAN"] <- "https://cloud.r-project.org/"
       options(repos=r)})

# #Modify Library Path
# #Note that only directories that exist at the time will be included.
# #It needs modified libPaths for accessing ALL Packages during New Startup. However...
# #Any change in libPaths at [1] earlier gets overwritten here [2]
if (!identical(.libPaths()[1], "C:/Softwares/R/LibFiles")) {
  .libPaths( c( "C:/Softwares/R/LibFiles", .libPaths()) )
}

# #To only download binaries and avoid 'latest code' type Errors in Windows.
options(pkgType = "binary")

# #Disable Scientific Notation because p-values, residuals etc. are irritating to convert
# #It might show many trailing zeroes but that can be handled separately.
#options(scipen = 0, digits = 7) #Default
options(scipen = 999)

# #Maximum Print Limit (Not lines, Characters. So, higher number is used.)
options(max.print=1000)

# #Set NPP as the external editor
options(editor = "C:/Program Files (x86)/Notepad++/notepad++.exe")

# #Change Prompt
options(prompt="> ")

# #Change the continuity prompt from (+), like (>), to wider (...) like Python
options(continue="... ")

# #Setup Max Width
options(width = 100)

# #Avoid GUI Dialogs in R (as implemented in RStudio)
options(menu.graphics=FALSE)

# #Form Feed Character (="\f" ="\014" ="^l") To suppress R Welcome Message (works for RStudio)
cat("\f")

# #Instruct R to, before anything else, echo a timestamp to the console and to '.Rhistory'
# #Load "utils" package for Sys.time()
library("utils")
.First <- function(){
  if(interactive()){
    message("# ### \"", getwd(), "\" ", Sys.time(), " #")
  }
}

# #tidyverse lifecycle package, to convert deprecated warnings to errors
options(lifecycle_verbosity = "error")

# #Warning: To upgrade Warning to Error for easy identification of problem chunk
# #Restore default: options(warn=0, error=NULL)
if(FALSE) options(warn = 2, error = recover)

```

### Hidden Namespace


```{r 'Z04-Hidden-Namespace', eval=FALSE}
#Hidden namespace to save objects & functions from "rm(list=ls())"
.z <- new.env()
attach(.z)

# #Object containing PATH to R Scripts
.z$RX <- paste0(getwd(), "/code/")

# #Object containing PATH to DATA
.z$XL <- paste0(getwd(), "/data/")

# #Object containing PATH to Images
.z$PX <- paste0(getwd(), "/images/")
```


## Warnings {.tabset .tabset-fade}

- "ForLater" Old Chunks and also not being evaluated

```{r 'Z04-Generate-Warning', eval=FALSE}
#jj <- try(as.numeric(c("5", "6", "seven")), silent = TRUE) #WARNING
```


### Generate {.unlisted .unnumbered}

```{r 'RS02B-Warning', eval=FALSE}
# #RS01RegEx
```

### Chunk Option {.unlisted .unnumbered}

```{r 'RS02D-UsingChunkOption', warning=TRUE, fig.cap="Show Warning using Chunk Option", eval=FALSE}
# #RS01RegEx
```

### suppressWarnings() {.unlisted .unnumbered}

```{r 'RS02E-Suppress', fig.cap="suppressWarnings()", eval=FALSE}
# #RS01RegEx
```

### tryCatch() {.unlisted .unnumbered}

```{r 'RS02F-TryCatchWarning', fig.cap="Show Warnings using tryCatch()", eval=FALSE}
# #RS01RegEx
```

### capture.output() {.unlisted .unnumbered}

```{r 'RS02G-CaptureOutput', fig.cap="Show Warnings using capture.output()", eval=FALSE}
# #RS01RegEx
```

### last.warning {.unlisted .unnumbered}

```{r 'RS02H-LastWarning', fig.cap="Show Warnings using last.warning", eval=FALSE}
# #RS01RegEx
```
### withCallingHandlers() {.unlisted .unnumbered}

```{r 'RS02I-withCallingHandlers', warning=TRUE, fig.cap="Show Warnings using withCallingHandlers()", eval=FALSE}
# #RS01RegEx
```

### options() {.unlisted .unnumbered}

```{r 'RS02J-WarningOption', eval=FALSE}
# #RS01RegEx
```

## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'Z04-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll, pkg, miss_pkg)
```

```{r 'Z04-Validation', include=FALSE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
