# Discrete Probability Distributions (BE05) {#be05}

```{r '305A', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## Overview

- This chapter covers 
  - [Discrete Probability](#discrete-prob-be05 "be05")

## Definitions

```{r '305D01', comment="", echo=FALSE, results='asis', cache=FALSE}
f_getDef("Discrete")
```

```{r '305D02', comment="", echo=FALSE, results='asis', cache=FALSE}
f_getDef("Continuous")
```

## Random Variable

```{definition 'Random-Variable'}
A \textcolor{pink}{random variable} is a numerical description of the outcome of an experiment. Random variables must assume numerical values. It can be either 'discrete' or 'continuous'.
```

```{definition 'Discrete-Random-Variable'}
A random variable that may assume either a finite number of values or an infinite sequence of values such as $0, 1, 2, \dots $ is referred to as a \textcolor{pink}{discrete random variable}. It includes factor type i.e. Male as 0, Female as 1 etc.
```

```{definition 'Continuous-Random-Variable'}
A random variable that may assume any numerical value in an interval or collection of intervals is called a \textcolor{pink}{continuous random variable}. It is given by \textcolor{pink}{$x \in [n, m]$}. If the entire line segment between the two points also represents possible values for the random variable, then the random variable is continuous.
```

## Developing Discrete Probability Distributions {#discrete-prob-be05}

```{definition 'Probability-Distribution'}
The \textcolor{pink}{probability distribution} for a random variable describes how probabilities are distributed over the values of the random variable.
```

```{definition 'Probability-Function'}
For a discrete random variable x, a \textcolor{pink}{probability function $f(x)$}, provides the probability for each value of the random variable.
```

- The use of the relative frequency method to develop discrete probability distributions leads to what is called an \textcolor{pink}{empirical discrete distribution}.
  - We treat the data as if they were the population and use the relative frequency method to assign probabilities to the experimental outcomes.

- Basic Requirements (Similar to the [Probability Basics](#assign-probability-be04 "be04"))
  1. \textcolor{pink}{$f(x) \geq 0$}
  1. \textcolor{pink}{$\sum {f(x)} = 1$}
- The simplest example of a discrete probability distribution given by a formula is the discrete uniform probability distribution; $f(x) = 1/n$, where $n$ is the number of values the random variable may assume

### Expected Value

```{definition 'Expected-Value-Discrete'}
The \textcolor{pink}{expected value, or mean}, of a random variable is a measure of the central location for the random variable. i.e. \textcolor{pink}{$E(x) = \mu = \sum xf(x)$}
```

- NOTE
  - The expected value is a weighted average of the values of the random variable where the weights are 
the probabilities.
  - The expected value does not have to be a value the random variable can assume. i.e. average need not to be integer
  
### Variance 

```{definition 'Variance-Discrete'}
The \textcolor{pink}{variance} is a weighted average of the squared deviations of a random variable from its mean. The weights are the probabilities. i.e. \textcolor{pink}{$\text{Var}(x) = \sigma^2 = \sum \{(x- \mu)^2 \cdot f(x)\}$}
```

## Bivariate Distributions

```{definition 'Bivariate'}
A probability distribution involving two random variables is called a \textcolor{pink}{bivariate probability distribution}. 
```

- NOTE:
  - Each outcome for a bivariate experiment consists of two values, one for each random variable. Example: Rolling a pair of dice

## Example-Dicarlo

```{r '303-Dicarlo', include=FALSE}
# #DicarloNY: Days with Number of Cars Sold per day for last 300 days
xxdicarlo <- tibble(Cars = 0:5, Days = c(54, 117, 72, 42, 12, 3))
bb <- xxdicarlo
bb <- bb %>% rename(x = Cars, Fx = Days) %>% mutate(across(Fx, ~./sum(Fx))) %>% 
  mutate(xFx = x * Fx, x_mu = x - sum(xFx), 
		     x_mu_sq = x_mu * x_mu, x_mu_sq_Fx = x_mu_sq * Fx) 
```

```{r '303-Dicarlo-KBL', echo=FALSE}
kbl(bb,
  caption = 'Variance Calculation',
  col.names = c("$x$", "$f(x)$", "$xf(x)$", "$(x - \\mu)$", "$(x - \\mu)^2$", "$\\sum {(x - \\mu)^{2}f(x)}$"),
  escape = FALSE, align = "c", booktabs = TRUE
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                html_font = "Consolas",	font_size = 12,
                full_width = FALSE,
                fixed_thead = TRUE
  ) %>%
# #Header Row Dark & Bold: RGB (48, 48, 48) =HEX (#303030)
	row_spec(0, color = "white", background = "#303030", bold = TRUE,
	         extra_css = "border-bottom: 1px solid; border-top: 1px solid"
	)
```


## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r '305-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk)
```

```{r '305-Validation', include=TRUE, cache=FALSE}
# #Summarised Packages and Objects
f_()
#
difftime(Sys.time(), k_start)
```

****
