# Discrete Probability Distributions (BE05) {#be05}

```{r '305A', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## Overview

- This chapter covers 
  - [Discrete Probability](#discrete-prob-be05 "be05")
  - [Bivariate Distributions](#bivariate-be05 "be05")
  - "ForLater" - Financial Applications, Exercises, Binomial, Poisson

## Definitions (Ref)

```{r '305D01', comment="", echo=FALSE, results='asis'}
f_getDef("Discrete")
```

```{r '305D02', comment="", echo=FALSE, results='asis'}
f_getDef("Continuous")
```

## Random Variable

```{definition 'Random-Variable'}
A \textcolor{pink}{random variable} is a numerical description of the outcome of an experiment. Random variables must assume numerical values. It can be either 'discrete' or 'continuous'.
```

```{definition 'Discrete-Random-Variable'}
A random variable that may assume either a finite number of values or an infinite sequence of values such as $0, 1, 2, \dots$ is referred to as a \textcolor{pink}{discrete random variable}. It includes factor type i.e. Male as 0, Female as 1 etc.
```

```{definition 'Continuous-Random-Variable'}
A random variable that may assume any numerical value in an interval or collection of intervals is called a \textcolor{pink}{continuous random variable}. It is given by \textcolor{pink}{$x \in [n, m]$}. If the entire line segment between the two points also represents possible values for the random variable, then the random variable is continuous.
```

## Discrete Probability Distributions {#discrete-prob-be05}

```{definition 'Probability-Distribution'}
The \textcolor{pink}{probability distribution} for a random variable describes how probabilities are distributed over the values of the random variable.
```

```{definition 'Probability-Function'}
For a discrete random variable x, a \textcolor{pink}{probability function $f(x)$}, provides the probability for each value of the random variable.
```

- The use of the relative frequency method to develop discrete probability distributions leads to what is called an \textcolor{pink}{empirical discrete distribution}.
  - We treat the data as if they were the population and use the relative frequency method to assign probabilities to the experimental outcomes.

- Basic Requirements (Similar to the [Probability Basics](#assign-probability-be04 "be04") & [Continuos Probability](#continuous-prob-be06 "be06"))
  1. \textcolor{pink}{$f(x) \geq 0$}
  1. \textcolor{pink}{$\sum {f(x)} = 1$}
- The simplest example of a discrete probability distribution given by a formula is the \textcolor{pink}{discrete uniform probability distribution}; $f(x) = 1/n$, where $n$ is the number of values the random variable may assume
  - Each possible value of the random variable has the same probability

### Expected Value

```{definition 'Expected-Value-Discrete'}
The \textcolor{pink}{expected value, or mean}, of a random variable is a measure of the central location for the random variable. i.e. \textcolor{pink}{$E(x) = \mu = \sum xf(x)$}
```

- NOTE
  - The expected value is a weighted average of the values of the random variable where the weights are 
the probabilities.
  - The expected value does not have to be a value the random variable can assume. i.e. average need not to be integer
  
### Variance 

```{definition 'Variance-Discrete'}
The \textcolor{pink}{variance} is a weighted average of the squared deviations of a random variable from its mean. The weights are the probabilities. i.e. \textcolor{pink}{$\text{Var}(x) = \sigma^2 = \sum \{(x- \mu)^2 \cdot f(x)\}$}
```

## Bivariate Distributions {#bivariate-be05}

```{definition 'Bivariate'}
A probability distribution involving two random variables is called a \textcolor{pink}{bivariate probability distribution}. A discrete bivariate probability distribution provides a probability for each pair of values that may occur for the two random variables.
```

- NOTE:
  - Each outcome for a bivariate experiment consists of two values, one for each random variable. Example: Rolling a pair of dice
  - Bivariate probabilities are often called \textcolor{pink}{joint probabilities}

## Ex Dicarlo {.tabset .tabset-fade}

```{r '305-Dicarlo', include=FALSE}
# #Dicarlo: Days with Number of Cars Sold per day for last 300 days
xxdicarlo <- tibble(Cars = 0:5, Days = c(54, 117, 72, 42, 12, 3))
#
bb <- xxdicarlo
bb <- bb %>% rename(x = Cars, Fx = Days) %>% mutate(across(Fx, ~./sum(Fx))) %>% 
  mutate(xFx = x * Fx, x_mu = x - sum(xFx), 
		     x_mu_sq = x_mu * x_mu, x_mu_sq_Fx = x_mu_sq * Fx) 
R_dicarlo_var_y_305 <- sum(bb$x_mu_sq_Fx)
# #Total Row
bb <- bb %>% 
  mutate(across(1, as.character)) %>% 
  add_row(summarise(., across(1, ~"Total")), summarise(., across(where(is.double), sum))) %>% 
  mutate(xFx = ifelse(x == "Total", paste0("mu = ", xFx), xFx),
         x_mu_sq_Fx = ifelse(x == "Total", paste0("sigma^2 = ", x_mu_sq_Fx), x_mu_sq_Fx)) %>% 
  mutate(across(4:5, ~ replace(., x == "Total", NA)))
```

### Table {.unlisted .unnumbered}

```{r '305-Dicarlo-KBL', echo=FALSE}
displ_names <- c("$x$", "$f(x)$", "$\\sum xf(x)$", "$(x - \\mu)$", "$(x - \\mu)^2$", 
                "$\\sum {(x - \\mu)^{2}f(x)}$") 
stopifnot(identical(ncol(bb), length(displ_names)))
#
kbl(bb,
  caption = 'Variance Calculation',
  col.names = displ_names,
  escape = FALSE, align = "c", booktabs = TRUE
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                html_font = "Consolas",	font_size = 12,
                full_width = FALSE,
				        #position = "float_left",
                fixed_thead = TRUE
  ) %>%
# #Header Row Dark & Bold: RGB (48, 48, 48) =HEX (#303030)
	row_spec(0, color = "white", background = "#303030", bold = TRUE,
	         extra_css = "border-bottom: 1px solid; border-top: 1px solid"
	)
```

### Data {.unlisted .unnumbered}

```{r '305-Dicarlo-A', ref.label=c('305-Dicarlo'), eval=FALSE}
#
```

### Change Class {.unlisted .unnumbered}

```{r '305-ColClass', eval=FALSE}
# #Change Column Classes as required
bb %>% mutate(across(1, as.character))
bb %>% mutate(across(everything(), as.character))
```

### Modify Value {.unlisted .unnumbered}

```{r '305-ModifyValue', eval=FALSE}
bb <- xxdicarlo
ii <- bb %>% rename(x = Cars, Fx = Days) %>% mutate(across(Fx, ~./sum(Fx))) %>% 
  mutate(xFx = x * Fx, x_mu = x - sum(xFx), 
		     x_mu_sq = x_mu * x_mu, x_mu_sq_Fx = x_mu_sq * Fx) 
# #Add Total Row
ii <- ii %>% 
  mutate(across(1, as.character)) %>% 
  add_row(summarise(., across(1, ~"Total")), summarise(., across(where(is.double), sum))) 
#
# #Modify Specific Row Values without using filter() 
# #filter() does not have 'un-filter()' function like group()-ungroup() combination
# #Selecting Row where x = "Total" and changing Column Values for Two Columns
ii <- ii %>% 
  mutate(xFx = ifelse(x == "Total", paste0("mu = ", xFx), xFx),
       x_mu_sq_Fx = ifelse(x == "Total", paste0("sigma^2 = ", x_mu_sq_Fx), x_mu_sq_Fx)) 
#
# #Selecting Row where x = "Total" and doing same replacement on Two Columns
ii %>% mutate(across(4:5, function(y) replace(y, x == "Total", NA)))
ii %>% mutate(across(4:5, ~ replace(., x == "Total", NA)))
```

## Ex Dicarlo GS {.tabset .tabset-fade}

```{r '305-DicarloGS', include=FALSE}
xxdicarlo_gs <- tibble(Geneva_Saratoga = c("x0", "x1", "x2", "x3"), 
             y0 = c(21, 21, 9, 3), y1 = c(30, 36, 42, 9), y2 = c(24, 33, 9, 6), 
             y3 = c(9, 18, 12, 3), y4 = c(2, 2, 3, 5), y5 = c(0, 1, 2, 0))
bb <- xxdicarlo_gs
#
# #Tibble Total SUM 
sum_bb <- bb %>% summarise(across(-1, sum)) %>% summarise(sum(.)) %>% pull(.)
#
# #Add Total Row and SUM Column
ii <- bb %>% 
  mutate(across(1, as.character)) %>% 
  add_row(summarise(., across(1, ~"Total")), summarise(., across(where(is.numeric), sum))) %>% 
  mutate(SUM = rowSums(across(where(is.numeric))))
#
# #Convert to Bivirate Probability Distribution and then add Total Row and SUM Column
jj <- bb %>% 
  mutate(across(where(is.numeric), ~./sum_bb)) %>% 
  mutate(across(1, as.character)) %>% 
  add_row(summarise(., across(1, ~"Total")), summarise(., across(where(is.numeric), sum))) %>% 
  mutate(SUM = rowSums(across(where(is.numeric)))) %>% 
  mutate(across(where(is.numeric), format, digits =1))
```

### Table {.unlisted .unnumbered}

```{r '305-DicarloGS-KBL', echo=FALSE}
bb <- ii
#displ_names <- c("") 
#stopifnot(identical(ncol(bb), length(displ_names)))
#
kk_ii <- kbl(bb,
  caption = 'Bivariate Table',
  #col.names = displ_names,
  escape = FALSE, align = "c", booktabs = TRUE
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                html_font = "Consolas",	font_size = 12,
                full_width = FALSE,
				        position = "float_left",
                fixed_thead = TRUE
  ) %>%
# #Header Row Dark & Bold: RGB (48, 48, 48) =HEX (#303030)
	row_spec(0, color = "white", background = "#303030", bold = TRUE,
	         extra_css = "border-bottom: 1px solid; border-top: 1px solid"
	)
#
bb <- jj
#displ_names <- c("") 
#stopifnot(identical(ncol(bb), length(displ_names)))
#
kk_jj <- kbl(bb,
  caption = 'Probability Distribution',
  #col.names = displ_names,
  escape = FALSE, align = "c", booktabs = TRUE
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                html_font = "Consolas",	font_size = 12,
                full_width = FALSE,
				        position = "float_left",
                fixed_thead = TRUE
  ) %>%
# #Header Row Dark & Bold: RGB (48, 48, 48) =HEX (#303030)
	row_spec(0, color = "white", background = "#303030", bold = TRUE,
	         extra_css = "border-bottom: 1px solid; border-top: 1px solid"
	)
# #Multiple Kable Tables
knitr::kables(list(kk_ii, kk_jj))
```

### DataGS {.unlisted .unnumbered}

```{r '305-DicarloGS-A', ref.label=c('305-DicarloGS'), eval=FALSE}
#
```

### Tibble Total SUM {.unlisted .unnumbered}

```{r '305-TibbleSum'}
bb <- xxdicarlo_gs
# #Assuming there is NO Total Column NOR Total Row and First Column is character
kk <- bb %>% summarise(across(where(is.numeric), sum)) %>% summarise(sum(.)) %>% pull(.)
ll <- bb %>% summarise(across(-1, sum)) %>% summarise(sum(.)) %>% pull(.)
stopifnot(identical(kk, ll))
print(kk)
```

### format() {.unlisted .unnumbered}

```{r '305-Digits'}
bb <- xxdicarlo_gs
# #Round off values to 1 significant digits i.e. 0.003 or 0.02
# #NOTE: This changes the column to "character"
bb %>% mutate(across(where(is.numeric), ~./sum_bb)) %>% 
  mutate(across(where(is.numeric), format, digits =1))
```

## Bivariate ... {.tabset .tabset-fade}

- Suppose we would like to know the probability distribution for total sales at both DiCarlo dealerships and the expected value and variance of total sales.
  - We can define $s = x + y$ as Total Sales.
  - Refer to the Tables \@ref(tab:305-DicarloGS-KBL) and \@ref(tab:305-Bivariate-KBL)
    - $f(s_0) = f(x_0, y_0) = 0.07$ 
    - $f(s_1) = f(x_0, y_1) + f(x_1, y_0) = 0.10 + 0.07 = 0.17$ 

```{r '305-BivariateVariance', include=FALSE}
bb <- xxdicarlo_gs
sum_bb <- bb %>% summarise(across(-1, sum)) %>% summarise(sum(.)) %>% pull(.)
# #Convert to Bivariate Probability Distribution
ii <- bb %>% mutate(across(where(is.numeric), ~./sum_bb)) %>% select(-1)
# #Using tapply(), sum the Matrix
jj <- tapply(X= as.matrix(ii), INDEX = LETTERS[row(ii) + col(ii)-1], FUN = sum)
# #Create Tibble
kk <- tibble(Fs = jj, ID = LETTERS[1:length(Fs)], s = 1:length(Fs) - 1) %>% 
  relocate(Fs, .after = last_col()) %>% 
  mutate(sFs = s * Fs, s_Es = s - sum(sFs), 
		     s_Es_sq = s_Es * s_Es, s_Es_sq_Fs = s_Es_sq * Fs) 
# #Save for Notebook
R_dicarlo_var_s_305 <- sum(kk$s_Es_sq_Fs)
# #For Printing
ll <- kk %>% 
  add_row(summarise(., across(1, ~"Total")), summarise(., across(where(is.double), sum))) %>% 
  mutate(across(where(is.numeric), format, digits =2)) %>% 
  mutate(sFs = ifelse(ID == "Total", paste0("E(s) = ", sFs), sFs),
         s_Es_sq_Fs = ifelse(ID == "Total", paste0("Var(s) = ", s_Es_sq_Fs), s_Es_sq_Fs)) %>% 
  mutate(across(c(2,5,6), ~ replace(., ID == "Total", NA)))
```

### Table {.unlisted .unnumbered}

```{r '305-Bivariate-KBL', echo=FALSE}
bb <- ll
displ_names <- c("$ID$", "$s$", "$f(s)$", "$\\sum sf(s)$", "$(s - E(s))$", "$(s - E(s))^2$", 
                "$\\sum {(s - E(s))^{2}f(s)}$") 
stopifnot(identical(ncol(bb), length(displ_names)))
#
kbl(bb,
  caption = 'Bivariate Expected Value and Variance',
  col.names = displ_names,
  escape = FALSE, align = "c", booktabs = TRUE
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                html_font = "Consolas",	font_size = 12,
                full_width = FALSE,
				        #position = "float_left",
                fixed_thead = TRUE
  ) %>%
# #Header Row Dark & Bold: RGB (48, 48, 48) =HEX (#303030)
	row_spec(0, color = "white", background = "#303030", bold = TRUE,
	         extra_css = "border-bottom: 1px solid; border-top: 1px solid"
	)
```

### Code {.unlisted .unnumbered}

```{r '305-BivariateVariance-A', ref.label=c('305-BivariateVariance'), eval=FALSE}
#
```

### Bivariate to Original {.unlisted .unnumbered}

```{r '305-DicarloG'}
bb <- xxdicarlo_gs
# #From the Bivariate get the original data
ii <- bb %>% 
  mutate(Fx = rowSums(across(where(is.numeric)))) %>% 
  select(1, 8) %>% 
  separate(col = Geneva_Saratoga, into = c(NA, "x"), sep = 1) %>% 
  mutate(across(1, as.integer))
# #Variance Calculation
jj <- ii %>% mutate(across(Fx, ~./sum(Fx))) %>% 
  mutate(xFx = x * Fx, x_mu = x - sum(xFx), 
		     x_mu_sq = x_mu * x_mu, x_mu_sq_Fx = x_mu_sq * Fx) 
# #Save for Notebook
R_dicarlo_var_x_305 <- sum(jj$x_mu_sq_Fx)
print(jj)
```

### Sum Diagonals {.unlisted .unnumbered}

```{r '303-SumDiagonal'}
bb <- xxdicarlo_gs
#
# #Tibble Total SUM 
sum_bb <- bb %>% summarise(across(-1, sum)) %>% summarise(sum(.)) %>% pull(.)
#
# #Convert to Bivirate Probability Distribution and Exclude First Character Column
ii <- bb %>% mutate(across(where(is.numeric), ~./sum_bb)) %>% select(-1)
#
# #(1A, 2B, 3C, 4D, 4E, 4F, 3G, 2H, 1I) 9 Unique Combinations = 24 (4x6) Experimental Outcomes 
matrix(data = LETTERS[row(ii) + col(ii)-1], nrow = 4)
# 
# #Using tapply(), sum the Matrix
jj <- tapply(X= as.matrix(ii), INDEX = LETTERS[row(ii) + col(ii)-1], FUN = sum)
print(jj)
# #In place of LETTERS, Numerical Index can also be used but Letters are more clear for grouping
#tapply(X= as.matrix(ii), INDEX = c(0:8)[row(ii) + col(ii)-1], FUN = sum)
#
# #Create Tibble
kk <- tibble(Fs = jj, ID = LETTERS[1:length(Fs)], s = 1:length(Fs) - 1) %>% 
  relocate(Fs, .after = last_col())
print(kk)
```

### String Split {.unlisted .unnumbered}

```{r '305-SplitString'}
bb <- xxdicarlo_gs
# #Separate String based on Position 
bb %>% separate(col = Geneva_Saratoga, into = c("A", "B"), sep = 1) 
```

## Covariance

- Covariance of random variables x and y is given by \textcolor{pink}{$\sigma_{xy}$}, Refer \@ref(eq:covariance-305)
  - NOTE: It does not look like \@ref(eq:covariance), but for now, I am assuming it is Equivalent
  - Calculated: $\text{Var}(s) = \text{Var}(x+y) =$ `r round(R_dicarlo_var_s_305, 3)`; $\text{Var}(y) =$ `r R_dicarlo_var_y_305`; $\text{Var}(x) =$ `r round(R_dicarlo_var_x_305, 3)`
  - Variance $\sigma_{xy} = \frac{2.3895 - 0.8696 - 1.25}{2} = 0.1350$
  - A covariance of .1350 indicates that daily sales at the two dealerships have a positive relationship. 

\begin{equation} 
     \sigma_{xy} = \frac{\text{Var}(x+y) - \text{Var}(x) - \text{Var}(y)}{2}
  (\#eq:covariance-305)
\end{equation} 

- Correlation of random variables x and y is given by, Refer \@ref(eq:correlation), \textcolor{pink}{$\rho_{xy}  = \frac{\sigma_{xy}}{\sigma_{x}\sigma_{y}}$} 
  - Where $\sigma_{x} = \sqrt{\text{Var}(x)} = \sqrt{0.8696} = 0.9325$; and $\sigma_{y} = \sqrt{\text{Var}(y)} = \sqrt{1.25} = 1.1180$
  - Correlation Coefficient $\rho_{xy}  = \frac{0.1350}{0.9325 \times 1.1180} = 0.1295$
  - The correlation coefficient of .1295 indicates there is a weak positive relationship between the random variables representing daily sales at the two dealerships. 

## Distributions

- "ForLater" 
  - \textcolor{pink}{Binomial Probability Distribution} - `dbinom(), pbinom(), qbinom(), rbinom()` 
    - It can be used to determine the probability of obtaining $x$ successes in $n$ trials.
    - 4 Assumptions must be TRUE
      1. The experiment consists of a sequence of $n$ identical trials.
      1. Two outcomes are possible on each trial, one called success and the other failure.
      1. The probability of a success $p$ does not change from trial to trial. Consequently, the probability of failure, $1 − p$, does not change from trial to trial.
      1. The trials are independent.
  - \textcolor{pink}{Poisson Probability Distribution} - `dpois(), ppois(), qpois(), rpois()` 
    - To determine the probability of obtaining $x$ occurrences over an interval of time or space.
    - 2 Assumptions must be TRUE
      1. The probability of an occurrence of the event is the same for any two intervals of equal length.
      1. The occurrence or nonoccurrence of the event in any interval is independent of the occurrence or nonoccurrence of the event in any other interval.
  - \textcolor{pink}{Hypergeometric Probability Distribution} 
    - Like the binomial, it is used to compute the probability of $x$ successes in $n$ trials. 
    - But, in contrast to the binomial, the probability of success changes from trial to trial.

## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r '305-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll, displ_names, kk_ii, kk_jj, sum_bb, xxdicarlo, xxdicarlo_gs, 
          R_dicarlo_var_s_305, R_dicarlo_var_x_305, R_dicarlo_var_y_305)
```

```{r '305-Validation', include=FALSE, cache=FALSE}
# #Summarised Packages and Objects
f_()
#
difftime(Sys.time(), k_start)
```

****
