# Interval Estimation {#c08}

```{r 'C08', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## Overview

- This chapter covers 
  - [Interval Estimation](#interval-c08 "c08") and [t-table](#get-t-c08 "c08")

## Interval Estimate {#interval-c08}

```{definition 'Interval-Estimate'}
Because a point estimator cannot be expected to provide the exact value of the population parameter, an \textcolor{pink}{interval estimate} is often computed by adding and subtracting a value, called the \textcolor{pink}{margin of error}, to the point estimate. $\text{Interval Estimate} = \text{Point Estimate} \pm \text{Margin of Error}$
```

```{definition 'Confidence-Interval'}
\textcolor{pink}{Confidence interval} is another name for an interval estimate. Normally it is given as \textcolor{pink}{$(1 - \alpha)$}. Ex: 95% confidence interval
```

```{definition 'Confidence-Coefficient'}
The confidence level expressed as a decimal value is the c\textcolor{pink}{onfidence coefficient}. i.e. 0.95 is the confidence coefficient for a 95% confidence level.
```


## Known SD {.unlisted .unnumbered .tabset .tabset-fade}

In order to develop an interval estimate of a population mean, either the population standard deviation $\sigma$ or the sample standard deviation $s$ must be used to compute the margin of error. In most applications $\sigma$ is not known, and $s$ is used to compute the margin of error. 

In some applications, large amounts of relevant historical data are available and can be used to estimate the population standard deviation prior to sampling. Also, in quality control applications where a process is assumed to be operating correctly, or 'in control,' it is appropriate to treat the population standard deviation as known. 

\textcolor{pink}{Sampling distribution of $\overline{x}$ can be used to compute the probability that $\overline{x}$ will be within a given distance of $\mu$.}

Example: Lloyd's Department Store

- Each week Lloyd's Department Store selects a simple random sample of 100 customers in order to learn about the amount spent per shopping trip. 
  - With $x$ representing the amount spent per shopping trip, the sample mean $\overline{x}$ provides a point estimate of $\mu$, the mean amount spent per shopping trip for the population of all Lloyd’s customers. Based on the historical data, Lloyd's now assumes a known value of $\sigma = 20$ for the population standard deviation. 
  - During the most recent week, Lloyd's surveyed 100 customers $(n = 100)$ and obtained a sample mean of $\overline{x} = 82$.
  - we can conclude that the sampling distribution of $\overline{x}$ follows a normal distribution with a standard error of $\sigma_{\overline{x}} = \frac{\sigma}{\sqrt{n}} = \frac{20}{\sqrt{100}} =2$. 
  - Because the sampling distribution shows how values of $\overline{x}$ are distributed around the population mean $\mu$, the sampling distribution of $\overline{x}$ provides information about the possible differences between $\overline{x}$ and $\mu$.
  - Using the standard normal probability table, we find that \textcolor{pink}{95\% of the values of any normally distributed random variable are within $\pm 1.96$ standard deviations of the mean i.e. $[\mu - 1.96 \sigma, \mu + 1.96\sigma]$}. 
    - Thus, 95% of the $\overline{x}$ values must be within $\pm 1.96 \sigma_{\overline{x}}$ of the mean $\mu$. 
    - In the Lloyd’s example we know that the sampling distribution of $\overline{x}$ is normally distributed with a standard error of $\sigma_{\overline{x}} =2$. 
    - we can conclude that 95% of all $\overline{x}$ values obtained using a sample size of $n = 100$ will be within $(\pm 1.96 \times 2 = \pm 3.92)$ of the population mean $\mu$.
  - As given above, sample mean was $\overline{x} = 82$
    - Interval estimate of $\overline{x} = 82 \pm 3.92 = [78.08, 85.92]$
    - Because 95% of all the intervals constructed using $\overline{x} = 82 \pm 3.92$ will contain the population mean, we say that we are 95% confident that the interval 78.08 to 85.92 includes the population mean $\mu$. 
    - We say that this interval has been established at the 95% confidence level. 
    - The value 0.95 is referred to as the \textcolor{pink}{confidence coefficient}, and the interval 78.08 to 85.92 is called the 95% \textcolor{pink}{confidence interval}.

Interval Estimate of a Population Mean: $\sigma$ known is given by equation \@ref(eq:interval-with-sigma)

\begin{equation}
  \begin{align}
    \overline{x} \pm z_{\alpha/2} \frac{\sigma}{\sqrt{n}}
  \end{align}
  (\#eq:interval-with-sigma)
\end{equation}

where $(1 − \alpha)$ is the confidence coefficient and $z_{\alpha/2}$ is the z-value providing an area of $\alpha/2$ in the upper tail of the standard normal probability distribution.

For a 95% confidence interval, the confidence coefficient is $(1 − \alpha) = 0.95$ and thus, $\alpha = 0.05$. Using the standard normal probability table, an area of $\alpha/2 = 0.05/2 = 0.025$ in the upper tail provides $z_{.025} = 1.96$. 

```{r '308-getZgivenAlpha'}
# #Find z-value for confidence interval 95% i.e. (1-alpha) = 0.95 i.e. alpha = 0.05
# #To look for Area under the curve towards Right only i.e. alpha/2 = 0.025
p_r_ii <- 0.025
p_l_ii <- 1 - p_r_ii
z_ii <- round(qnorm(p = p_l_ii, lower.tail = TRUE), 4)
cat(paste0("(Left) P(z) = ", format(p_l_ii, nsmall = 3), " (i.e. (Right) 1-P(z) = ", 
           format(p_r_ii, nsmall = 3), ") at z = ", z_ii, "\n"))
```

"ForLater" - Various Common alpha values and z-values Table - (Book Table 8.1)

## Unknown SD {.tabset .tabset-fade}


```{definition 't-distribution'}
When $s$ is used to estimate $\sigma$, the margin of error and the interval estimate for the population mean are based on a probability distribution known as the \textcolor{pink}{t distribution}.
```

The t distribution is a family of similar probability distributions, with a specific t distribution depending on a parameter known as the\textcolor{pink}{degrees of freedom}. As the number of degrees of freedom increases, the difference between the t distribution and the standard normal distribution becomes smaller and smaller. 

Just as $z_{0.025}$ was used to indicate the z value providing a 0.025 area in the upper tail of a standard normal distribution, $t_{0.025}$ indicates a 0.025 area in the upper tail of a t distribution. In general, the notation $t_{\alpha/2}$ represents a t value with an area of $\alpha/2$ in the upper tail of the t distribution.

As the degrees of freedom increase, the t distribution approaches the standard normal distribution. Ex:  $t_{0.025} = 2.262 \,(\text{DOF} = 9)$, $t_{0.025} = 2.200 \,(\text{DOF} = 60)$, and $t_{0.025} = 1.96 \,(\text{DOF} = \infty) = z_{0.025}$

Interval Estimate of a Population Mean: $\sigma$ Unknown is given by equation \@ref(eq:interval-no-sigma)

\begin{equation}
  \begin{align}
    \overline{x} \pm t_{\alpha/2} \frac{s}{\sqrt{n}}
  \end{align}
  (\#eq:interval-no-sigma)
\end{equation}

where $s$ is the sample standard deviation, $(1 − \alpha)$ is the confidence coefficient and $t_{\alpha/2}$ is the t-value providing an area of $\alpha/2$ in the upper tail of the t distribution with $n-1$ degrees of freedom.

Refer equation \@ref(eq:sd), the expression for the sample standard deviation is 

\begin{equation*} 
    {s} = \sqrt{\frac{\sum \left(x_i - \bar{x}\right)^2}{n-1}}
\end{equation*} 

Why $(n-1)$ are the degrees of freedom

- Degrees of freedom refer to the number of independent pieces of information that go into the computation. i.e. $\{(x_{1}-\overline{x}), (x_{2}-\overline{x}), \ldots, (x_{n}-\overline{x})\}$
- However, $\sum (x_{i}-\overline{x}) =0$ for any data set. 
- Thus, only $(n − 1)$ of the $(x_{i}-\overline{x})$ values are independent.
  - if we know $(n − 1)$ of the values, the remaining value can be determined exactly by using the condition.

\textcolor{pink}{Larger sample sizes are needed if the distribution of the population is highly skewed or includes outliers.}

### Cal T  {#get-t-c08 .unlisted .unnumbered}

```{r '308-getT'}
# #Find t-value for confidence interval 95% i.e. (1-alpha) = 0.95 i.e. alpha = 0.05
# #To look for Area under the curve towards Right only i.e. alpha/2 = 0.025
p_r_ii <- 0.025
p_l_ii <- 1 - p_r_ii
#
# #t-tables are unique for different degrees of freedom i.e. for DOF = 9 
dof_ii <- 9
t_ii <- round(qt(p = p_l_ii, df = dof_ii, lower.tail = TRUE), 4)
cat(paste0("(Left) P(t) = ", format(p_l_ii, nsmall = 3), " (i.e. (Right) 1-P(t) = ", 
           format(p_r_ii, nsmall = 3), ") at t = ", t_ii, " (dof = ", dof_ii, ")\n"))
```
 
### qt() {.unlisted .unnumbered}

```{r '308-getTgivenAlpha'}
# #Find t-value for confidence interval 95% i.e. (1-alpha) = 0.95 i.e. alpha = 0.05
# #To look for Area under the curve towards Right only i.e. alpha/2 = 0.025
p_r_ii <- 0.025
p_l_ii <- 1 - p_r_ii
#
# #t-tables are unique for different degrees of freedom i.e. for DOF = 9 
dof_ii <- 9
t_ii <- round(qt(p = p_l_ii, df = dof_ii, lower.tail = TRUE), 4)
cat(paste0("(Left) P(t) = ", format(p_l_ii, nsmall = 3), " (i.e. (Right) 1-P(t) = ", 
           format(p_r_ii, nsmall = 3), ") at t = ", t_ii, " (dof = ", dof_ii, ")\n"))
#
dof_ii <- 60
t_ii <- round(qt(p = p_l_ii, df = dof_ii, lower.tail = TRUE), 4)
cat(paste0("(Left) P(t) = ", format(p_l_ii, nsmall = 3), " (i.e. (Right) 1-P(t) = ", 
           format(p_r_ii, nsmall = 3), ") at t = ", t_ii, " (dof = ", dof_ii, ")\n"))
#
dof_ii <- 600
t_ii <- round(qt(p = p_l_ii, df = dof_ii, lower.tail = TRUE), 4)
cat(paste0("(Left) P(t) = ", format(p_l_ii, nsmall = 3), " (i.e. (Right) 1-P(t) = ", 
           format(p_r_ii, nsmall = 3), ") at t = ", t_ii, " (dof = ", dof_ii, ")\n"))
#
# #t-table have Infinity Row which is same as z-table. For DOF >100, it can be used.
dof_ii <- Inf
t_ii <- round(qt(p = p_l_ii, df = dof_ii, lower.tail = TRUE), 4)
cat(paste0("(Left) P(t) = ", format(p_l_ii, nsmall = 3), " (i.e. (Right) 1-P(t) = ", 
           format(p_r_ii, nsmall = 3), ") at t = ", t_ii, " (dof = ", dof_ii, ")\n"))

#
z_ii <- round(qnorm(p = p_l_ii, lower.tail = TRUE), 4)
cat(paste0("(Left) P(z) = ", format(p_l_ii, nsmall = 3), " (i.e. (Right) 1-P(z) = ", 
           format(p_r_ii, nsmall = 3), ") at z = ", z_ii, "\n"))
```

### Ex: Credit Card {.unlisted .unnumbered .tabset .tabset-fade}

```{r '308-CreditCards'}
# #A sample of n = 70 households provided the credit card balances.
xxCreditCards <- c(9430, 7535, 4078, 5604, 5179, 4416, 10676, 1627, 10112, 6567, 13627, 18719, 14661, 12195, 10544, 13659, 7061, 6245, 13021, 9719, 2200, 10746, 12744, 5742, 7159, 8137, 9467, 12595, 7917, 11346, 12806, 4972, 11356, 7117, 9465, 19263, 9071, 3603, 16804, 13479, 14044, 6817, 6845, 10493, 615, 13627, 12557, 6232, 9691, 11448, 8279, 5649, 11298, 4353, 3467, 6191, 12851, 5337, 8372, 7445, 11032, 6525, 5239, 6195, 12584, 15415, 15917, 12591, 9743, 10324)
f_setRDS(xxCreditCards)
```

```{r '308-Credit-Estimate-Interval'}
bb <- f_getRDS(xxCreditCards)
mean_bb <- mean(bb)
sd_bb <- sd(bb)
dof_bb <- length(bb) - 1L
# #t-value for confidence interval 95% | (1-alpha) = 0.95 | alpha = 0.05 | alpha/2 = 0.025
p_r_ii <- 0.025
p_l_ii <- 1 - p_r_ii
#
dof_ii <- dof_bb
t_ii <- round(qt(p = p_l_ii, df = dof_ii, lower.tail = TRUE), 4)
cat(paste0("(Left) P(t) = ", format(p_l_ii, nsmall = 3), " (i.e. (Right) 1-P(t) = ", 
           format(p_r_ii, nsmall = 3), ") at t = ", t_ii, " (dof = ", dof_ii, ")\n"))
#
# #Interval Estimate
err_margin_bb <- t_ii * sd_bb / sqrt(length(bb))
est_l <- mean_bb - err_margin_bb
est_r <- mean_bb + err_margin_bb
#
cat(paste0("Normal Sample (n=", length(bb), ", mean=", mean_bb, ", sd=", round(sd_bb, 1),
           "):\n Point Estimate = ", mean_bb, ", Margin of error = ", round(err_margin_bb, 1), 
           ", ", (1-2*p_r_ii) * 100, "% confidence interval is [", 
           round(est_l, 1), ", ", round(est_r, 1), "]"))
```

## "ForLater" {.unlisted .unnumbered}

- Determining the Sample Size
- Population Proportion

## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'C08-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll, dof_bb, dof_ii, err_margin_bb, est_l, est_r, mean_bb, p_l_ii, p_r_ii, sd_bb, t_ii, xxCreditCards, z_ii)
```

```{r 'C08-Validation', include=FALSE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
