# Two Populations {#c10}

```{r 'C10', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## Overview

- "Inference About Means and Proportions with Two Populations"
  - "ForLater"


## Summary

How interval estimates and hypothesis tests can be developed for situations involving two populations when the difference between the two population means or the two population proportions is of prime importance.

Example 

- To develop an interval estimate of the difference between the mean starting salary for a population of men and the mean starting salary for a population of women.
- To conduct a hypothesis test to determine whether any difference is present between the proportion of defective parts in a population of parts produced by supplier A and the proportion of defective parts in a population of parts produced by supplier B. 

- The matched sample design is generally preferred to the independent sample design because the matched-sample procedure often improves the precision of the estimate.


## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'C10-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll)
```

```{r 'C10-Validation', include=FALSE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
