# Simple Linear Regression {#c14}

```{r 'C14', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## Overview

- 
  - "ForLater" - Everything


## Simple Linear Regression Model {.unlisted .unnumbered .tabset .tabset-fade}

```{definition 'Regression-Analysis'}
\textcolor{pink}{Regression analysis} can be used to develop an equation showing how two or more variables variables are related.
```


```{definition 'Variable-Dependent'}
The variable being predicted is called the \textcolor{pink}{dependent variable $({y})$}.
```


```{definition 'Variable-Independent'}
The variable or variables being used to predict the value of the dependent variable are called the \textcolor{pink}{independent variables $({x})$}.
```


```{definition 'Simple-Linear-Regression'}
The simplest type of regression analysis involving one independent variable and one dependent variable in which the relationship between the variables is approximated by a straight line, is called \textcolor{pink}{simple linear regression}.
```


```{definition 'Regression-Model'}
The equation that describes how ${y}$ is related to ${x}$ and an error term is called the \textcolor{pink}{regression model}. For example, simple linear regression model is given by equation \@ref(eq:simple-linear) 
```

\begin{equation}
  {y} = {\beta}_0 + {\beta}_1 {x} + {\epsilon}
  (\#eq:simple-linear)
\end{equation}

Note

- ${\beta}_0$ and ${\beta}_1$ are referred to as the \textcolor{pink}{parameters of the model}
- The random variable, \textcolor{pink}{error term $({\epsilon})$}, accounts for the variability in ${y}$ that cannot be explained by the linear relationship between ${x}$ and ${y}$.




## Summary


## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'C14-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll)
```

```{r 'C14-Validation', include=FALSE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
