# Introduction to Data Mining {#c31}
> Definitions and Exercises are from the Book [@Larose]

```{r 'C31', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```

## Overview

- "An Introduction to Data Mining and Predictive Analytics"

```{definition 'Data-Mining-331'}
\textcolor{pink}{Data mining} is the process of discovering useful patterns and trends in large data sets. 
```

```{definition 'Predictive-Analytics-331'}
\textcolor{pink}{Predictive analytics} is the process of extracting information from large data sets in order to make predictions and estimates about future outcomes.
```

- The Cross-Industry Standard Process for Data Mining (CRISP-DM) (Iterative)
  - Business/Research Understanding Phase
    - Clearly enunciate the project objectives and requirements in terms of the business or research unit as a whole. 
    - Translate these goals and restrictions into the formulation of a data mining problem definition.
    - Prepare a preliminary strategy for achieving these objectives.
  - Data Understanding Phase 
    - Collect the data.
    - Use \textcolor{pink}{Exploratory Data Analysis (EDA)} to familiarize yourself with the data, and discover initial insights. 
    - Evaluate the quality of the data. 
    - if desired, select interesting subsets that may contain actionable patterns. 
  - Data Preparation Phase 
    - Raw | Select | Filter | Subset | Clean | Transform 
  - Modeling Phase
    - Select and apply appropriate modeling techniques
    - Calibrate
  - Evaluation Phase
    - Models must be evaluated for quality and effectiveness. 
    - Also, determine whether the model in fact achieves the objectives set for it
    - Establish whether some important facet of the business or research problem has not been sufficiently accounted for. 
  - Deployment Phase / Report / Publish

## Data Mining Methods {#mining-def-c31}

- [Data Mining Methods](#mining-b19 "b19") and [Definitions](#mining-def-c31 "c31")
  - Data mining methods may be categorized as either supervised or unsupervised.
  - Most data mining methods are supervised methods.
  - Unsupervised : Clustering, PCA, Factor Analysis, Association Rules, RFM
  - Supervised : 
    - Regression (Continuous Target) : Linear Regression, Regularised Regression, Decision Tree, Random Forest
    - Classification (Categorical Target) : Decision Trees, Random Forest, Logistic Regression, k-nearest neighbors, Bagging and Boosting, Neural Networks

```{definition 'Description'}
\textcolor{pink}{Description} of patterns and trends often suggest possible explanations for existence of theme within the data.
```

```{definition 'Estimation'}
In \textcolor{pink}{estimation}, we approximate the value of a numeric target variable using a set of numeric and/or categorical predictor variables. Methods: Point Estimation, Confidence Interval Estimation, Simple Linear Regression, Correlation, Multiple Regression etc.
```

```{definition 'Prediction'}
\textcolor{pink}{Prediction} is similar to classification and estimation, except that for prediction, the results lie in the future. Estimation methods are also used for Prediction. Additional Methods: k-nearest neighbor methods, decision trees, neural networks etc.
```

```{definition 'Classification'}
\textcolor{pink}{Classification} is similar to estimation, however, instead of approximating the value of a numeric target variable, the target variable is categorical. 
```

```{r 'C31D01', comment="", echo=FALSE, results='asis'}
f_getDef("Clustering")
```

```{r 'C31D02', comment="", echo=FALSE, results='asis'}
f_getDef("Cluster")
```

```{r 'C31D03', comment="", echo=FALSE, results='asis'}
f_getDef("Affinity-Analysis")
```


## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'C31-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll)
```

```{r 'C31-Validation', include=FALSE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
