# Dimension Reduction {#c34}

```{r 'C34', include=FALSE, cache=FALSE}
sys.source(paste0(.z$RX, "A99Knitr", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "000Packages", ".R"), envir = knitr::knit_global())
sys.source(paste0(.z$RX, "A00AllUDF", ".R"), envir = knitr::knit_global())
#invisible(lapply(f_getPathR(A09isPrime), knitr::read_chunk))
```


## Dimensions {.tabset .tabset-fade}

> 10 or 11 Dimensionsions are enough for the Universe. How many are needed for your data!


### Multicollinearity {.unlisted .unnumbered}

```{definition 'Multicollinearity'}
\textcolor{pink}{Multicollinearity} is a condition where some of the predictor variables are strongly correlated with each other. 
```

- Problems: Multicollinearity
  - Multicollinearity leads to instability in the solution space, leading to possible incoherent results, such as in multiple regression, where a multicollinear set of predictors can result in a regression which is significant overall, even when none of the individual variables is significant. 
  - Even if such instability is avoided, inclusion of variables which are highly correlated tends to overemphasize a particular component of the model, as the component is essentially being double counted.
- Problems: Too many variables
  - The sample size needed to fit a multivariate function grows exponentially with the number of variables.
  - The use of too many predictor variables to model a relationship with a response variable can unnecessarily complicate the interpretation of the analysis, and violates the principle of parsimony
    - i.e. keep the number of predictors to such a size that would be easily interpreted. 
  - Also, retaining too many variables may lead to overfitting
    - i.e. generality of the findings is hindered because new data do not behave the same as the training data for all the variables.

### Parsimony {.unlisted .unnumbered}

```{definition 'Principle-of-Parsimony'}
\textcolor{pink}{Principle of parsimony} is the problem-solving principle that "entities should not be multiplied beyond necessity". 
```

- It is inaccurately paraphrased as "the simplest explanation is usually the best one". 
- It advocates that when presented with competing hypotheses about the same prediction, one should select the solution with the fewest assumptions, and that this is not meant to be a way of choosing between hypotheses that make different predictions. 

### Overfitting & Underfitting {.unlisted .unnumbered}

```{definition 'Overfitting'}
\textcolor{pink}{Overfitting} is the production of an analysis that corresponds too closely to a particular set of data, and may therefore fail to fit additional data or predict future observations reliably.
```


```{definition 'Underfitting'}
\textcolor{pink}{Underfitting} occurs when a statistical model cannot adequately capture the underlying structure of the data. 
```

- An overfitted model is a statistical model that contains more parameters than can be justified by the data.
  - The essence of overfitting is to have unknowingly extracted some of the residual variation (i.e., the noise) as if that variation represented underlying model structure.
  - over-fitting occurs when a model begins to "memorize" training data rather than "learning" to generalize from a trend.
- Under-fitting would occur, for example, when fitting a linear model to non-linear data. Such a model will tend to have poor predictive performance.
- Overfitting Example
  - If the number of parameters is the same as or greater than the number of observations, then a model can perfectly predict the training data simply by memorizing the data in its entirety. Such a model, though, will typically fail severely when making predictions.
  - A noisy linear dataset can be fitted by a polynomial function also which would give a perfect fit. However, the linear function (in this case) would be better in extrapolating beyond the fitted data.
- To decrease the chance or amount of overfitting
  - model comparison, cross-validation, regularization, early stopping, pruning, Bayesian priors, or dropout
  - To explicitly penalize overly complex models 
  - To evaluate the model performance on a set of data not used for training, which is assumed to approximate the typical unseen data that a model will encounter. 

## Dimension-reduction methods 

- These use the correlation structure among the predictor variables to accomplish the following: 
  - To reduce the number of predictor items. 
  - To help ensure that these predictor items are independent. 
  - To provide a framework for interpretability of the results. 

- Dimension-reduction methods: 
  - PCA
  - Factor analysis 
  - User-defined composites 
  
## PCA

```{definition 'PCA'}
\textcolor{pink}{Principal components analysis (PCA)} seeks to explain the correlation structure of a set of predictor variables ${m}$, using a smaller set of linear combinations of these variables, called components ${k}$. PCA acts solely on the predictor variables, and ignores the target variable.
```


- Suppose that the original variables ${\{X_1, X_2, \ldots, X_m\}}$ form a coordinate system in m-dimensional space. 
  - Let each variable ${X_i}$ represent an ${n \times 1}$ vector, where ${n}$ is number of records.
  - The principal components represent a new coordinate system, found by rotating the original system along the directions of maximum variability.
- Analysis
  - Standardize the data, so that the mean for each variable is zero, and the standard deviation is one.
  - ${X_i \to Z_i = \frac{X_i - {\mu}_i}{{\sigma}_{ii}}}$
  - The covariance is a measure of the degree to which two variables vary together.
  - A positive covariance indicates that, when one variable increases, the other tends to increase, while a negative covariance indicates that, when one variable increases, the other tends to decrease. 
  - ${{\sigma}_{ii}^2}$ denotes the variance of ${X_i}$. 
    - If ${X_i}$ and ${X_j}$ are independent, then ${{\sigma}_{ij}^2 = 0}$; but reverse may not be TRUE i.e. ${{\sigma}_{ij}^2 = 0}$ does not imply that ${X_i}$ and ${X_j}$ are independent. 
    - Note that the covariance measure is not scaled, so that changing the units of measure would change the value of the covariance.
    - The correlation coefficient ${r_{ij}} = \frac{{\sigma}_{ij}^2}{{\sigma}_{ii}{\sigma}_{jj}}$ avoids this difficulty by scaling the covariance by each of the standard deviations.
    - Then, the correlation matrix is denoted as ${\rho}$


```{definition 'Eigenvalues'}
Let \textcolor{pink}{B} be an $m \times m$ matrix, and let \textcolor{pink}{I} be the $m \times m$ identity matrix. Then the scalars $\{\lambda_1, \lambda_2, \ldots, \lambda_m\}$ are said to be the \textcolor{pink}{eigenvalues of B} if they satisfy $|B - \lambda I| = 0$, where $|Q|$ denotes the determinant of Q.
```


```{definition 'Eigenvector'}
Let \textcolor{pink}{B} be an $m \times m$ matrix, and let $\lambda$ be an eigenvalue of B. Then nonzero $m \times 1$ vector \textcolor{pink}{e} is said to be an \textcolor{pink}{eigenvector of B}, if $B e = 𝜆e$.
```

- The total variability in the standardized set of predictors equals the sum of the variances of the Z-vectors, which equals the sum of the variances of the components, which equals the sum of the eigenvalues, which equals the numer of predictors
  - $\sum_{i=1}^m {\text{Var}({Y_i})} = \sum_{i=1}^m {\text{Var}({Z_i})} = \sum_{i=1}^m {\lambda_i} = m$
- The partial correlation between a given component and a given predictor variable is a function of an eigenvector and an eigenvalue. Specifically, $\text{Corr}(Y_i, Z_j) = e_{ij}\sqrt{\lambda_i}$, where $\{ (\lambda_1, e_1), (\lambda_2, e_2), \ldots, (\lambda_m, e_m)\}$ are the eigenvalue–eigenvector pairs for the correlation matrix $\rho$, and we note that $\lambda_1 \geq \lambda_2 \geq \ldots \geq \lambda_m$. In other words, the eigenvalues are ordered by size. (A partial correlation coefficient is a correlation coefficient that takes into account the effect of all the other variables.)
- The proportion of the total variability in Z that is explained by the $i^{\text{th}}$ principal component is the ratio of the $i^{\text{th}}$ eigenvalue to the number of variables, that is, the ratio $\frac{\lambda_i}{m}$.

## Data Housing

\textcolor{pink}{Please import the "C34-cadata.txt".} 

- Source: http://lib.stat.cmu.edu/datasets/houses.zip
- About: [20640, 9]
  - It provides census information from all the block groups from the 1990 California census. 
  - For this data set, a block group has an average of 1425.5 people living in an area that is geographically compact. 
  - Block groups were excluded that contained zero entries for any of the variables. 
  - Variables: median house value (Target), median income, housing median age, total rooms, total bedrooms, population, households, latitude, and longitude.


```{r 'C34-Housing', include=FALSE, eval=FALSE}
# #read_table() can handle double space delimited file.
tbl <- read_table(paste0(.z$XL, "C34-cadata.txt"), skip = 27, 
                col_names = c("median_house_value", "median_income", "housing_median_age", 
                              "total_rooms", "total_bedrooms", "population", "households", 
                              "latitude", "longitude"))
stopifnot(!"problems" %in% names(attributes(tbl)))
#problems(tbl)
#attr(tbl, "problems") <- NULL
attr(tbl, "spec") <- NULL
xxC34Housing <- tbl
f_setRDS(xxC34Housing)
```

```{r 'C34-GetHousing', include=FALSE}
bb <- aa <- xxC34Housing <- f_getRDS(xxC34Housing)
```

## Partition Train & Test Data

```{r 'C34-Partition'}
set.seed(3)
# #sample() and its varaints can be used for Partitioning of Dataset
if(FALSE) {
  # #This approach is difficult to extend for 3 or more splits
  # #Further, Using floor() multiple times might result in loss of a row 
  train_idx <- sample.int(n = nrow(bb), size = floor(0.8 * nrow(bb)), replace = FALSE)
  train_idx <- sample(seq_len(nrow(bb)), size = floor(0.8 * nrow(bb)), replace = FALSE)  
  train_bb <- bb[train_idx, ]
  test_bb <- bb[-train_idx, ]
}
#
# #For 3 or more splits
#brk_bb = c(train = 0.8, test = 0.1, validate = 0.1)
brk_bb = c(train = 0.9, test = 0.1)
idx_bb = sample(cut(seq_len(nrow(bb)), nrow(bb) * cumsum(c(0, brk_bb)), labels = names(brk_bb)))
#
# #Splits by Fixed Numbers not Percentages
if(FALSE) {
  brk_bb = c(train = 18570, test = nrow(bb))
  idx_bb = sample(cut(seq_len(nrow(bb)), c(0, brk_bb), labels = names(brk_bb)))
}
# #List of Multiple Tibbles
aa = split(bb, idx_bb)
#
# #nrow(), ncol(), dim() can be applied
vapply(aa, nrow, FUN.VALUE = integer(1))
stopifnot(identical(nrow(bb), sum(vapply(aa, nrow, FUN.VALUE = integer(1)))))
```


## Basics, Skewness & Normality



WORKING HERE



```{r 'C34-Normality', include=FALSE}
bb <- aa$train
#str(bb)
ii <- bb %>%  
  pivot_longer(everything(), names_to = "Predictors", values_to = "Values") %>% 
  mutate(across(Predictors, factor, levels = unique(Predictors)))
#
jj <- ii %>% group_by(Predictors) %>% 
  summarise(Min = min(Values), Max = max(Values), SD = sd(Values), 
            Mean = mean(Values), Median = median(Values), Mode = f_getMode(Values),
            Unique = length(unique(Values)), isNA = sum(is.na(Values)), 
            Skewness = skewness(Values), 
            p_Shapiro = shapiro.test(Values)$p.value, 
            isNormal = ifelse(p_Shapiro > 0.05, TRUE, FALSE))
#
kk <- jj %>% 
  mutate(across(where(is.numeric), round, digits = 4)) %>% 
  mutate(across(where(is.numeric), format, digits = 3, nsmall = 0, 
                replace.zero = TRUE, zero.print = "0", scientific = FALSE, drop0trailing = TRUE)) 
```

```{r 'C33T01', echo=FALSE}
kbl(kk,
  caption = "(C33T01) Churn: Normality",
  #col.names = displ_names,
  escape = FALSE, align = "c", booktabs = TRUE
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"),
                html_font = "Consolas",	font_size = 12,
                full_width = FALSE,
                #position = "float_left",
                fixed_thead = TRUE
  ) %>%
# #Header Row Dark & Bold: RGB (48, 48, 48) =HEX (#303030)
	row_spec(0, color = "white", background = "#303030", bold = TRUE,
	         extra_css = "border-bottom: 1px solid; border-top: 1px solid"
	) %>% 
# #Conditional Row Text Highlight. Background etc. can also be done.
  row_spec(which(!kk$isNormal), color = "red") # %>% row_spec(which(kk$isNormal), color = "black")
```

## Validation {.unlisted .unnumbered .tabset .tabset-fade}

```{r 'C34-Cleanup', include=FALSE, cache=FALSE}
f_rmExist(aa, bb, ii, jj, kk, ll, c32churn, xxB16Cars, xxB18Churn)
```

```{r 'C34-Validation', include=TRUE, cache=FALSE}
# #SUMMARISED Packages and Objects (BOOK CHECK)
f_()
#
difftime(Sys.time(), k_start)
```

****
